### 2018-11-15 paper/paper.tex
abstract draft

% This project is part of the Tensor Estimator project
% Copyright 2018 the authors.

\documentclass[modern]{aastex62}

% aastex parameters
%%\hypersetup{linkcolor=red,citecolor=green,filecolor=cyan,urlcolor=magenta}
\received{not yet but hopefully soon!!}
%\revised{not yet}
%\accepted{not yet}
%\submitjournal{ApJ}
\shorttitle{tensor estimator for large-scale structure}
\shortauthors{storey-fisher and hogg}

% typography

% affiliations
\newcommand{\ccpp}{\affiliation{%
    Center for Cosmology and Particle Physics,
    Department of Physics,
    New York University}}
\newcommand{\flatiron}{\affiliation{%
    Flatiron Institute, Simons Foundation}}
\newcommand{\cds}{\affiliation{%
    Center for Data Science,
    New York University}}
\newcommand{\mpia}{\affiliation{%
    Max-Planck-Institut f\"{u}r Astronomie, Heidelberg}}

\begin{document}\sloppy\sloppypar\raggedbottom\frenchspacing

\title{A Vectorized Correlation Function Estimator For Large-Scale Structure}
%\title{A Generalized Correlation Function Estimator for Galaxy Surveys}

%Tensorial estimator?
%Generalized estimator?
%Tensorial estimator? 
%Projected estimator?
%Continuous estimator?
%Vecstimator?

\author{Kate Storey-Fisher}
\ccpp

\author[0000-0003-2866-9403]{David W. Hogg}
\ccpp
\flatiron
\cds
\mpia

\begin{abstract}\noindent
% Context
The two-point correlation function (2PCF) is one of the most commonly used statistics to characterize clustering in galaxy surveys. However, current estimators of the 2PCF have significant limitations. The standard Landy-Szalay estimator evaluates the 2PCF in bins of radial separation between objects, and the choice of bins introduces a trade-off between bias and variance. Further trade-offs are required to determine the dependence of clustering on galaxy properties, typically by subsampling the objects. A marked correlation function, which weights the 2PCF based on some property, avoids some of these issues but still requires binning and contains a bias with sample selection.
% Aims
We present a new, vectorized estimator for the 2PCF that resolves these issues. 
% Methods
It projects pairs of objects onto any set of continuous basis functions, eliminating the need for binning. These functions can also depend on galaxy properties, so the estimator effectively computes a regression against these properties. This produces a vector that encodes the clustering information, and the 2PCF can be quickly evaluated at any value of the given property, giving a precise, unbiased characterization of the evolution.
% Results
We apply this estimator to the main galaxy sample of the Sloan Digital Sky Survey and demonstrate that it precisely characterizes the luminosity evolution of clustering. We also use the estimator to directly fit for the baryon acoustic feature and show that it extracts Fisher-optimal statistics. These results indicate that this new method provides the accuracy and precision needed to analyze upcoming galaxy surveys.

\end{abstract}

\section{Introduction}

Large-scale structure (LSS) is well-described by $\Lambda$CDM.

LSS is also used to constrain galaxy formation.

The Landy-Szalay estimator is traditionally used to characterize clustering in the LSS.

The marked correlation function has been used to understand the clustering dependence on galaxy properties.

The 2PCF is closely related to the power spectrum, but these are computed separately.

This is a probabilistic/frequentist approach.

\section{Generalities}

The Landy-Szalay estimator is justified by\ldots

Estimating clustering is closely related to least-squares fitting.

From these, we can infer the form of the estimator.

\section{Methods}

We derive/define this new estimator.

We show that it reduces to the Landy-Szalay estimator.

We prove that the estimator is invariant under affine transformations.

These show that the estimator is correct.

We implement this estimator using \ldots

\section{Tests and Results}

To test this estimator, we generate a Gaussian random field with a known power spectrum (and thus known correlation function). We sample the field to obtain a set of tracer objects.

We apply the estimator to the SDSS main galaxy sample. \ldots We reproduce the results of Zehavi et al, showing that the clustering shows a strong dependence on luminosity. \ldots We find that we can more precisely track this luminosity evolution.

We use this estimator to linearize around the BAO feature, using an LRG dataset. Show Fisher-optimal! (Different Fisher.)

\section{Discussion}

Covariance matrix

We estimated the 2PCF with fewer components to achieve the same level of detail.

In the future, we could simultaneously estimate the correlation function and the power spectrum.

We could also investigate: gradients, anomalies, growth of structure, directly estimating the cosmological parameters, calibrating for systematics

We have reformulated the correlation function in terms of pairs of objects.

Our performance is by definition the same as traditional estimators due to the limiting factor of pair-finding.

Use our estimator it is the best!

\acknowledgements
It is a pleasure to thank\ldots

\end{document}

### 2018-11-15 paper/paper.tex
small abstract edits

@@
 %\submitjournal{ApJ}
-\shorttitle{tensor estimator for large-scale structure}
+\shorttitle{vectorized estimator for large-scale structure}
 \shortauthors{storey-fisher and hogg}
@@
 % Methods
-It projects pairs of objects onto any set of continuous basis functions, eliminating the need for binning.
+It generalizes the Landy-Szalay estimator to depend on a tensor of random catalog pairs.
+The estimator then projects pairs onto any set of continuous basis functions, eliminating the need for binning.
 These functions can also depend on galaxy properties, so the estimator effectively computes a regression against these properties.
-This produces a vector that encodes the clustering information, and the 2PCF can be quickly evaluated at any value of the given property, giving a precise, unbiased characterization of the evolution.
+This produces a vector that encodes the clustering information, and the 2PCF can be quickly evaluated at any value of the given property, giving a precise, unbiased, and smooth characterization of the evolution.
 % Results

### 2018-11-20 paper/paper.tex
todo lists, abstract

--- 2018-11-15 paper/paper.tex
@@
 % Copyright 2018 the authors.
+
+% to-do
+% -----
+% - get yourself an ORCID and use it.
+% - Is ``vectorized'' the right word? It's what I say, but now I'm doubting it. We could go with linear-regression. Or affine-invariant. Or unbinned. Or continuous-function?
+
+% style notes
+% -----------
+% - line break at sentence breaks? Or more frequently? So git diff works good?
 
@@
 %\submitjournal{ApJ}
-\shorttitle{tensor estimator for large-scale structure}
+\shorttitle{vectorized correlation function estimator}
 \shortauthors{storey-fisher and hogg}
@@
 \ccpp
-\flatiron
 \cds
 \mpia
+\flatiron
 
@@
 % Context
-The two-point correlation function (2PCF) is one of the most commonly used statistics to characterize clustering in galaxy surveys.
+The two-point correlation function (2PCF) the most important statistic in structure formation.
 However, current estimators of the 2PCF have significant limitations.
-The standard Landy-Szalay estimator evaluates the 2PCF in bins of radial separation between objects, and the choice of bins introduces a trade-off between bias and variance.
-Further trade-offs are required to determine the dependence of clustering on galaxy properties, typically by subsampling the objects.
-A marked correlation function, which weights the 2PCF based on some property, avoids some of these issues but still requires binning and contains a bias with sample selection.
+The standard Landy-Szalay (LS) estimator evaluates the 2PCF in bins of radial separation between objects, and the choice of bins introduces a trade-off between bias and variance.
+Further trade-offs are required to determine the dependence of clustering on galaxy properties or redshift, typically by subsampling the objects.
 % Aims
-We present a new, vectorized estimator for the 2PCF that resolves these issues. 
+We present a new, vectorized estimator (Foo) for the 2PCF that obviates binning in separation or anything else.
+Instead of estimating the 2PCF in bins, it estimates the parameters of a continuous-function representation of the 2PCF in the form of a linear superposition of basis functions.
 % Methods
-It projects pairs of objects onto any set of continuous basis functions, eliminating the need for binning.
-These functions can also depend on galaxy properties, so the estimator effectively computes a regression against these properties.
-This produces a vector that encodes the clustering information, and the 2PCF can be quickly evaluated at any value of the given property, giving a precise, unbiased characterization of the evolution.
+Inspired by linear least-squares fitting, the Foo estimator marks the numerator terms of LS with vectors of basis functions of pair properties, and the denominator with a tensor of outer products of those functions.
+The function estimates the best-fit linear combination of basis functions to describe the 2PCF, in the limit that LS would be applicable.
 % Results
-We apply this estimator to the main galaxy sample of the Sloan Digital Sky Survey and demonstrate that it precisely characterizes the luminosity evolution of clustering.
-We also use the estimator to directly fit for the baryon acoustic feature and show that it extracts Fisher-optimal statistics.
-These results indicate that this new method provides the accuracy and precision needed to analyze upcoming galaxy surveys.
-
+We apply Foo to artificial data to demonstrate that it can determine the clustering in a representation that provides more accuracy with fewer basis functions than LS, thus reducing requirements on mocks for covariance estimation.
+We apply Foo to the Main Sample of the Sloan Digital Sky Survey (SDSS) to show that Foo can measure the luminosity-dependence of clustering without binning samples in luminosity.
+We apply Foo to the Luminous Red Galaxy Sample of SDSS-III (DO WE?) to show that Foo can extract a small set of Fisher-optimal statistics for measuring the location of the Baryon Acoustic Feature.
+We discuss other applications and limitations of Foo for present and future studies of large-scale structure; in principle it can be used to unify real-space and Fourier-space approaches.
 \end{abstract}

### 2018-11-20 paper/paper.tex
small wording improvements

@@
 % Context
-The two-point correlation function (2PCF) the most important statistic in structure formation.
+The two-point correlation function (2PCF) is the most important statistic in structure formation.
 However, current estimators of the 2PCF have significant limitations.
 The standard Landy-Szalay (LS) estimator evaluates the 2PCF in bins of radial separation between objects, and the choice of bins introduces a trade-off between bias and variance.
-Further trade-offs are required to determine the dependence of clustering on galaxy properties or redshift, typically by subsampling the objects.
+Further binning trade-offs are required to determine the dependence of clustering on galaxy properties or redshift.
 % Aims
@@
 Inspired by linear least-squares fitting, the Foo estimator marks the numerator terms of LS with vectors of basis functions of pair properties, and the denominator with a tensor of outer products of those functions.
-The function estimates the best-fit linear combination of basis functions to describe the 2PCF, in the limit that LS would be applicable.
+The function estimates the best-fit linear combination of basis functions to describe the 2PCF, in the same limit in which LS is justifiably optimal for binned applications.
 % Results
@@
 We apply Foo to the Luminous Red Galaxy Sample of SDSS-III (DO WE?) to show that Foo can extract a small set of Fisher-optimal statistics for measuring the location of the Baryon Acoustic Feature.
-We discuss other applications and limitations of Foo for present and future studies of large-scale structure; in principle it can be used to unify real-space and Fourier-space approaches.
+We discuss other applications and limitations of Foo for present and future studies of large-scale structure; in principle it can be used to unify real-space and Fourier-space approaches because it can be used to estimate parameters for any wavelet or Fourier basis.
 \end{abstract}
@@
 \section{Introduction}
+
+Binning is sinning!
 

### 2018-11-20 paper/paper.tex
tweaks

@@
 
-\title{A Vectorized Correlation Function Estimator For Large-Scale Structure}
+\title{\textbf{No More Bins:\\ A Vectorized Correlation Function Estimator For Large-Scale Structure}}
 %\title{A Generalized Correlation Function Estimator for Galaxy Surveys}
@@
 We apply Foo to the Main Sample of the Sloan Digital Sky Survey (SDSS) to show that Foo can measure the luminosity-dependence of clustering without binning samples in luminosity.
-We apply Foo to the Luminous Red Galaxy Sample of SDSS-III (DO WE?) to show that Foo can extract a small set of Fisher-optimal statistics for measuring the location of the Baryon Acoustic Feature.
+We apply Foo to the Luminous Red Galaxy Sample of SDSS-III (DO WE?) to show that Foo can extract a compact set of Fisher-optimal statistics for measuring the location of the baryon acoustic feature.
 We discuss other applications and limitations of Foo for present and future studies of large-scale structure; in principle it can be used to unify real-space and Fourier-space approaches because it can be used to estimate parameters for any wavelet or Fourier basis.

### 2019-03-18 paper/paper.tex
update abstract

--- 2018-11-20 paper/paper.tex
@@
 
-\title{A Vectorized Correlation Function Estimator For Large-Scale Structure}
+\title{\textbf{A Continuous Correlation Function Estimator for Large-Scale Structure}}
+%\title{\textbf{A Continuous Representation of the Correlation Function Estimator for Large-Scale Structure}}
+%\title{\textbf{Correlation Function Estimation with Continuous Functions for Large-Scale Structure}}
+%\title{\textbf{Estimating the Two-Point Correlation Function with Continuous Functions}}
+%\title{\textbf{Projecting Two-Point Correlations onto General Basis Functions: A New Estimator For Large-Scale Structure}}
+%\title{\textbf{A Correlation Function Estimator with General Basis Functions for Large-Scale Structure}}
+%\title{\textbf{A Generalized Correlation Function Estimator For Large-Scale Structure}}
+%\title{\textbf{No More Bins:\\ A Vectorized Correlation Function Estimator For Large-Scale Structure}}
 %\title{A Generalized Correlation Function Estimator for Galaxy Surveys}
@@
 %Continuous estimator?
-%Vecstimator?
 
@@
 The two-point correlation function (2PCF) is the most important statistic in structure formation.
-However, current estimators of the 2PCF have significant limitations.
+Current estimators of the 2PCF, which characterizes the clustering of density field tracers, have significant limitations.
 The standard Landy-Szalay (LS) estimator evaluates the 2PCF in bins of radial separation between objects, and the choice of bins introduces a trade-off between bias and variance.
-Further binning trade-offs are required to determine the dependence of clustering on galaxy properties or redshift.
 % Aims
-We present a new, vectorized estimator (Foo) for the 2PCF that obviates binning in separation or anything else.
-Instead of estimating the 2PCF in bins, it estimates the parameters of a continuous-function representation of the 2PCF in the form of a linear superposition of basis functions.
-% Methods
-Inspired by linear least-squares fitting, the Foo estimator marks the numerator terms of LS with vectors of basis functions of pair properties, and the denominator with a tensor of outer products of those functions.
-The function estimates the best-fit linear combination of basis functions to describe the 2PCF, in the same limit in which LS is justifiably optimal for binned applications.
+We present a new, generalized estimator (Foo) for the 2PCF that obviates binning in separation or any other property.
 % Results
-We apply Foo to artificial data to demonstrate that it can determine the clustering in a representation that provides more accuracy with fewer basis functions than LS, thus reducing requirements on mocks for covariance estimation.
-We apply Foo to the Main Sample of the Sloan Digital Sky Survey (SDSS) to show that Foo can measure the luminosity-dependence of clustering without binning samples in luminosity.
-We apply Foo to the Luminous Red Galaxy Sample of SDSS-III (DO WE?) to show that Foo can extract a small set of Fisher-optimal statistics for measuring the location of the Baryon Acoustic Feature.
-We discuss other applications and limitations of Foo for present and future studies of large-scale structure; in principle it can be used to unify real-space and Fourier-space approaches because it can be used to estimate parameters for any wavelet or Fourier basis.
+The Foo estimator replaces the binned pair counts of LS with a continuous-function representation in the form of a linear superposition of basis functions.
+These functions can take advantage of the known form of the 2PCF, and can depend on other properties of the pairs in addition to their separation.
+Inspired by linear least-squares fitting, Foo marks the numerator terms of LS with vectors of these functions, and the denominator with a tensor of their outer products.
+The estimator outputs the best-fit linear combination of basis functions to describe the 2PCF, in the same limit in which LS is justifiably optimal for binned applications. 
+% Results
+We apply Foo to the Luminous Red Galaxy Sample of the Sloan Digital Sky Survey (SDSS) to show that it estimates the clustering in a representation that provides more accuracy with fewer basis functions than LS, thus reducing requirements on mocks for covariance estimation.
+We demonstrate that Foo can extract a compact set of Fisher-optimal statistics for measuring the location of the baryon acoustic feature.
+We discuss other applications and limitations of Foo for present and future studies of large-scale structure; it could be used to determine the dependence of clustering on galaxy properties such as luminosity, and in principle it could unify real-space and Fourier-space approaches because by using a Fourier basis representation.
 \end{abstract}

### 2019-03-21 paper/paper.tex
first draft of intro

@@
-% This project is part of the Tensor Estimator project
-% Copyright 2018 the authors.
+% This project is part of the Continuous-Function Estimator project
+% Copyright 2019 the authors.
 
@@
 
-\documentclass[modern]{aastex62}
+\documentclass[twocolumn]{aastex62}
+
+\usepackage[sort&compress]{natbib}
+\usepackage{xspace}
 
@@
 %%\hypersetup{linkcolor=red,citecolor=green,filecolor=cyan,urlcolor=magenta}
-\received{not yet but hopefully soon!!}
+\received{XXX}
 %\revised{not yet}
-%\accepted{not yet}
+\accepted{YYY}
 %\submitjournal{ApJ}
-\shorttitle{vectorized correlation function estimator}
-\shortauthors{storey-fisher and hogg}
+\shorttitle{A Continuous Correlation Function Estimator}
+\shortauthors{Storey-Fisher and Hogg}
 
 % typography
+\newcommand{\cf}{2pcf\xspace} %2pF? 2PCF? %TODO: fix spacing after
+\newcommand{\Est}{The Continuous-Function Estimator\xspace}
+\newcommand{\est}{the Continuous-Function Estimator\xspace}
 
@@
 
-\title{\textbf{A Continuous Correlation Function Estimator for Large-Scale Structure}}
+\title{\textbf{A Continuous-Function Estimator of the Correlation Function for Large-Scale Structure}}
+%\title{\textbf{A Continuous Correlation Function Estimator for Large-Scale Structure}}
+%\title{\textbf{A Continuous-Function Correlation Function Estimator for Large-Scale Structure}}
+%\title{\textbf{Projecting the Correlation Function onto Continuous Functions for Large-Scale Structure Surveys}
 %\title{\textbf{A Continuous Representation of the Correlation Function Estimator for Large-Scale Structure}}
@@
 
-%Tensorial estimator?
-%Generalized estimator?
-%Tensorial estimator? 
-%Projected estimator?
-%Continuous estimator?
+%Generalized Estimator?
+%Projected Estimator?
+%Continuous Estimator?
+%Continuous-Function Estimator?
+%Linear-Regression Estimator?
+%Vectorized estimator?
 
-\author{Kate Storey-Fisher}
+\author[0000-0001-8764-7103]{Kate Storey-Fisher}
 \ccpp
@@
 % Context
-The two-point correlation function (2PCF) is the most important statistic in structure formation.
-Current estimators of the 2PCF, which characterizes the clustering of density field tracers, have significant limitations.
-The standard Landy-Szalay (LS) estimator evaluates the 2PCF in bins of radial separation between objects, and the choice of bins introduces a trade-off between bias and variance.
+The two-point correlation function (\cf) is the most important statistic in structure formation.
+Current estimators of the \cf, used to measure the clustering of density field tracers, have significant limitations.
+The standard Landy-Szalay (LS) estimator evaluates the \cf in bins of radial separation between objects, and the choice of bins introduces a trade-off between bias and variance.
 % Aims
-We present a new, generalized estimator (Foo) for the 2PCF that obviates binning in separation or any other property.
+We present a new, generalized estimator, \est, for the \cf that obviates binning in separation or any other property.
 % Results
-The Foo estimator replaces the binned pair counts of LS with a continuous-function representation in the form of a linear superposition of basis functions.
-These functions can take advantage of the known form of the 2PCF, and can depend on other properties of the pairs in addition to their separation.
-Inspired by linear least-squares fitting, Foo marks the numerator terms of LS with vectors of these functions, and the denominator with a tensor of their outer products.
-The estimator outputs the best-fit linear combination of basis functions to describe the 2PCF, in the same limit in which LS is justifiably optimal for binned applications. 
+\Est replaces the binned pair counts of LS with a continuous-function representation in the form of a linear superposition of basis functions.
+These functions can take advantage of the known form of the \cf, and can depend on other properties of the pairs in addition to their separation.
+Inspired by linear least-squares fitting, \est marks the numerator terms of LS with vectors of these functions, and the denominator with a tensor of their outer products.
+The estimator outputs the best-fit linear combination of basis functions to describe the \cf, in the same limit in which LS is justifiably optimal for binned applications. 
 % Results
-We apply Foo to the Luminous Red Galaxy Sample of the Sloan Digital Sky Survey (SDSS) to show that it estimates the clustering in a representation that provides more accuracy with fewer basis functions than LS, thus reducing requirements on mocks for covariance estimation.
-We demonstrate that Foo can extract a compact set of Fisher-optimal statistics for measuring the location of the baryon acoustic feature.
-We discuss other applications and limitations of Foo for present and future studies of large-scale structure; it could be used to determine the dependence of clustering on galaxy properties such as luminosity, and in principle it could unify real-space and Fourier-space approaches because by using a Fourier basis representation.
+We show that it can estimate the clustering of artificial data in representations that provide more accuracy with fewer basis functions than LS, thus reducing requirements on mocks for covariance estimation.
+We apply \est to the Luminous Red Galaxy sample of the Sloan Digital Sky Survey (SDSS), and show can extract a compact set of Fisher-optimal statistics for measuring the location of the baryon acoustic feature.
+We discuss other applications and limitations of \est for present and future studies of large-scale structure, including determining the dependence of clustering on galaxy properties and potentially unifying real-space and Fourier-space approaches by using a Fourier basis representation.
 \end{abstract}
+
+\keywords{cosmology: large-scale structure --- galaxies: statistics} %?
 
@@
 
-Binning is sinning!
+The large-scale structure (LSS) of the universe is a critical probe of fundamental cosmology. 
+It encodes information about the physics of the early universe and the subsequent expansion history.
+In particular, the LSS provides a probe of the Baryon Acoustic Oscillations (BAO), density fluctuations resulting from baryon-photon coupling in the early universe.
+The distance traveled by these waves imprints a feature on the statistical description of the LSS, which can be used to determine the characteristic BAO length scale \citep{EisensteinHu1998}.
+The LSS also contains the signature of redshift-space distortions caused by the peculiar velocities of galaxies, which probe the growth rate of structure \citep{Kaiser1987}.
+Additionally, the LSS can be used to constrain galaxy formation in conjunction with models of galaxy bias (e.g. \citealt{Hamilton1988}). %?
+With current observations, the LSS is well-described by a cold dark matter model with a cosmological constant, the standard $\Lambda$CDM model.
+Upcoming galaxy surveys will observe larger volumes with improved measurements, allowing us to test $\Lambda$CDM to even higher precision.
 
-Large-scale structure (LSS) is well-described by $\Lambda$CDM.
+We characterize the LSS by using luminous sources to trace the underlying matter density field.
+These tracers are often to taken to be galaxies, but can also be galaxy clusters, quasars and other sources.
+% should i say that hereafter we will take them to be galaxies?
+The clustering of these objects is measured with two-point statistics, namely the power spectrum $P(k)$ and the two-point correlation function (\cf). %name xi_r?
+These characterize the clustering in Fourier space and real space, respectively, with the \cf defined as the Fourier Transform of the power spectrum; in principle, they contain the same information.
+%show eqn?
+However, in practical applications the survey boundaries introduce nontrivial issues in computing these statistics, leading to diverging approaches to their computation with a significant difference in expense.
+The \cf requires more computational power and extra survey products, but it is an incredibly useful tool; for instance, it well-suited to the analysis of the BAO feature which manifests at a single scale in real space.
 
-LSS is also used to constrain galaxy formation.
+The \cf measures the excess probability that any two galaxies are separated by a given distance, compared to a uniform distribution; effectively, it characterizes the strength of clustering at a given spatial scale.
+% need to discuss statistical homogeneity & isotropy? (ergodicity?)
+In calculating the \cf, the boundaries of the surveys prevent us from directly summing pair counts due to nontrivial edge effects.
+To account for the survey boundaries as well as corrupted regions due to issues such as bright foreground stars, a set of random points are Poisson-distributed within the acceptable survey window. 
+The pairwise correlations of these points, which by construction are unclustered, are used to normalize out the survey window when estimating the \cf of the clustered data.
+Typically, this requires random points on the order of 10-100 times the number of data points, making the random correlations the limiting factor in \cf computation.
 
-The Landy-Szalay estimator is traditionally used to characterize clustering in the LSS.
+The \cf is computed in bins of radial separation, meaning that in practice it measures the volume average of the \cf over the bin.
+The choice of bins requires a trade-off between bias and variance: fewer bins may bias the result, while more bins increases the variance of measurement.
+Finite-width bins also result in a loss of information about the property in which one is binning. 
+Especially as we strive to extreme precision in large-scale structure analyses, we should be maximizing the information we extract from the data.
+Moreover, we note that standard estimators are fundamentally biased due to this finite binning; we will address this in a future work. %TODO: figure out this wording
 
-The marked correlation function has been used to understand the clustering dependence on galaxy properties.
+%some papers have equations (even this eqn) in intro - but should bump it to motivation and just explain?
+The current standard estimator was proposed by \cite{LandySzalay1993}, hereafter LS, which is based on summing all data pairs $DD$ with a given separation and using data-random pairs $DR$ and random pairs $RR$ to correct for the survey boundary.
+The correlation function $\xi_k(r)$ for the $k^\mathrm{th}$ separation bin is
+%assume we should keep this notation even though it's clunky, as it's recognizable
+%should we use estimator notation with a hat on xi?
+\begin{equation}
+\xi_k(r) = \frac{DD_k(r) - 2DR_k(r) + RR_k(r)}{RR_k(r)}.
+\end{equation}
+Compared with other estimators based on combinations of $DD$, $DR$ and $RR$, LS has been shown to have the lowest bias and variance \citep{Kerscher2000}.
+%is this necessary? not sure where it fits
+Estimators of the \cf must also take into account the imperfect nature of the survey area that is in the window, including the target completeness and fiber collisions; typically each galaxy pair is assigned a weight based on these.
 
-The 2PCF is closely related to the power spectrum, but these are computed separately.
+Variations on the random catalog pair count method have been proposed in recent years.
+\cite{Demina2016} replaced the $DR$ and $RR$ terms with an integral over the probability map, reducing computation time and increasing precision.
+An estimator proposed by \cite{VargasMagana2013} iterates over sets of mock catalogs to find an optimal linear combination of data and random pair counts, reducing the bias and variance.
+The marked correlation function \citep{WhitePadmanabhan2009} sums weights based on properties of the galaxies one is interested in, such as the local density or galaxy luminosity, and avoids the use of a random catalog.
+The estimators described so far have all taken probabilistic approaches; some have also taken a likelihood approach.
+\cite{BaxterRozo2013} introduced a maximum likelihood estimator for the \cf, which achieves lower variance compared to the LS estimator, enabling finer binning and requiring a smaller random catalog for the same precision.
 
-This is a probabilistic/frequentist approach.
+These estimators present improvements to LS, but they all still require binning in separation.
+Some require additional computational costs or layers of complexity, so the standard formulation of LS continues to be the default estimator used in analyses.
 
-\section{Generalities}
+In this paper, we present a new estimator for the correlation function, \est, which generalizes the LS estimator to produce a continuous estimation of the \cf. 
+\Est projects the galaxy pairs onto a set of continuous basis functions and computes the best-fit linear combination of these functions.
+The basis representation can depend on the pair separation as well as other desired properties, and can also utilize the known form of the \cf.
+For top-hat basis functions, \est exactly reduces to the LS estimator. 
+This estimator removes the need for binning and allows for the \cf to be represented by fewer basis functions, requiring fewer mock catalogs to compute the covariance matrix.
+It is particularly well-suited to the analysis of LSS features such as the BAO peak; we find that we can more accurately locate the peak with fewer components.
+
+%spell out acronyms again that were defined in the abstract?
+This paper is organized as follows.
+In Section \ref{sec:motiv}, we motivate and derive our estimator.
+In Section \ref{sec:est}, we prove its correctness and describe our implementation, and demonstrate its application on a simulated data set.
+Our results on the SDSS LRG sample are shown in Section \ref{sec:app}.
+We discuss the implications and other possible applications in Section \ref{sec:discuss}.
+
+\section{Motivation} \label{sec:motiv}
 
@@
 
-\section{Methods}
+% TODO: fix this capitalization issue
+\section{\Est} \label{sec:est}
 
@@
 
+We show that in the limit of infinitesimal bins\ldots
+
 These show that the estimator is correct.
 
-We implement this estimator using \ldots
+We show that it extracts Fisher-optimal information.
 
-\section{Tests and Results}
+We implement this estimator based on the correlation function package \texttt{corrfunc} by \citep{citeXXX} \ldots
+The code is open-source and available at XXX.
 
-To test this estimator, we generate a Gaussian random field with a known power spectrum (and thus known correlation function). We sample the field to obtain a set of tracer objects.
+We show its performance on a simulated Gaussian random field.
 
-We apply the estimator to the SDSS main galaxy sample.
-\ldots We reproduce the results of Zehavi et al, showing that the clustering shows a strong dependence on luminosity.
-\ldots We find that we can more precisely track this luminosity evolution.
+\section{Application to survey data} \label{sec:app}
 
-We use this estimator to linearize around the BAO feature, using an LRG dataset. Show Fisher-optimal! (Different Fisher.)
+We apply the estimator to the SDSS DR3 sample from \cite{Eisenstein2005}.
 
-\section{Discussion}
+The reconstructed dataset we use is described in \cite{Kazin2010}.
 
-Covariance matrix
+We use \est to linearize around the BAO feature.
 
-We estimated the 2PCF with fewer components to achieve the same level of detail.
+\section{Discussion} \label{sec:discuss}
+
+We estimated the \cf with fewer components to achieve the same level of precision.
+
+We note that our estimator doesn't rely on LS; it could be implemented for various linear combination of pair counts, given certain properties. %aka the tensor term
+
+Our performance is by definition the same (marginally more) compared to traditional estimators due to the limiting factor of pair-finding.
+
+We discuss the impact on computing the covariance matrix.
+
+We discuss the relation of our estimator to the optimized estimator of \cite{VargasMagana2013}, which also uses a linear combination approach.
 
@@
 
-We could also investigate: gradients, anomalies, growth of structure, directly estimating the cosmological parameters, calibrating for systematics
-
-We have reformulated the correlation function in terms of pairs of objects.
-
-Our performance is by definition the same as traditional estimators due to the limiting factor of pair-finding.
-
-Use our estimator it is the best!
+We could also investigate: gradients, anomalies, growth of structure, directly estimating the cosmological parameters, calibrating for systematics, clustering dependence galaxy properties
 
@@
 
+%need style file error
+%\bibliographystyle{apj} 
+\bibliography{paper}
+
 \end{document}

### 2019-06-21 paper/paper.tex
worked on motivation/methods, appendix

@@
 
-\documentclass[twocolumn]{aastex62}
+%\documentclass[twocolumn]{aastex62}
+%\documentclass[12 pt]{article}
+\documentclass[modern]{aastex62}
 
@@
 
-% typography
+% language
 \newcommand{\cf}{2pcf\xspace} %2pF? 2PCF? %TODO: fix spacing after
@@
 \newcommand{\est}{the Continuous-Function Estimator\xspace}
+\newcommand{\LS}{LS\xspace}
+\newcommand{\foreign}[1]{\textsl{#1}}
+\newcommand{\etc}{\foreign{etc}}
+
+% math
+\newcommand{\inv}{^{-1}}
+\newcommand{\T}{^{\mathsf{T}}}
+
+% margins
+%\addtolength{\topmargin}{-0.75in}
+%\addtolength{\textheight}{1.50in}
 
@@
 \begin{document}\sloppy\sloppypar\raggedbottom\frenchspacing
-
-\title{\textbf{A Continuous-Function Estimator of the Correlation Function for Large-Scale Structure}}
+\title{\textbf{Two-point statistics without bins: A correlation function estimator for large-scale structure in a continuous-function basis}}
+%\title{\textbf{A Continuous-Function Estimator of the Correlation Function for Large-Scale Structure}}
 %\title{\textbf{A Continuous Correlation Function Estimator for Large-Scale Structure}}
@@
 Current estimators of the \cf, used to measure the clustering of density field tracers, have significant limitations.
-The standard Landy-Szalay (LS) estimator evaluates the \cf in bins of radial separation between objects, and the choice of bins introduces a trade-off between bias and variance.
+The standard Landy-Szalay (\LS) estimator evaluates the \cf in bins of radial separation between objects, and the choice of bins introduces a trade-off between bias and variance.
 % Aims
@@
 % Results
-\Est replaces the binned pair counts of LS with a continuous-function representation in the form of a linear superposition of basis functions.
+\Est replaces the binned pair counts of \LS with a continuous-function representation in the form of a linear superposition of basis functions.
 These functions can take advantage of the known form of the \cf, and can depend on other properties of the pairs in addition to their separation.
-Inspired by linear least-squares fitting, \est marks the numerator terms of LS with vectors of these functions, and the denominator with a tensor of their outer products.
-The estimator outputs the best-fit linear combination of basis functions to describe the \cf, in the same limit in which LS is justifiably optimal for binned applications. 
+Inspired by linear least-squares fitting, \est marks the numerator terms of \LS with vectors of these functions, and the denominator with a tensor of their outer products.
+The estimator outputs the best-fit linear combination of basis functions to describe the \cf, in the same limit in which \LS is justifiably optimal for binned applications. 
 % Results
-We show that it can estimate the clustering of artificial data in representations that provide more accuracy with fewer basis functions than LS, thus reducing requirements on mocks for covariance estimation.
-We apply \est to the Luminous Red Galaxy sample of the Sloan Digital Sky Survey (SDSS), and show can extract a compact set of Fisher-optimal statistics for measuring the location of the baryon acoustic feature.
+We demonstrate that it can estimate the clustering of artificial data in representations that provide more accuracy with fewer basis functions than \LS, thus reducing requirements on mocks for covariance estimation.
+We apply \est to the Luminous Red Galaxy sample of the Sloan Digital Sky Survey (SDSS), and show that it can extract a compact set of Fisher-optimal statistics for measuring the location of the baryon acoustic feature.
 We discuss other applications and limitations of \est for present and future studies of large-scale structure, including determining the dependence of clustering on galaxy properties and potentially unifying real-space and Fourier-space approaches by using a Fourier basis representation.
@@
 
-The \cf is computed in bins of radial separation, meaning that in practice it measures the volume average of the \cf over the bin.
+The \cf is computed in bins of radial separation, meaning that in practice it measures the volume average of the \cf over the bin. %is this accurate?
 The choice of bins requires a trade-off between bias and variance: fewer bins may bias the result, while more bins increases the variance of measurement.
 Finite-width bins also result in a loss of information about the property in which one is binning. 
-Especially as we strive to extreme precision in large-scale structure analyses, we should be maximizing the information we extract from the data.
-Moreover, we note that standard estimators are fundamentally biased due to this finite binning; we will address this in a future work. %TODO: figure out this wording
+Especially as we strive for extreme precision in large-scale structure analyses, we should be maximizing the information we extract from the data.
+Moreover, we note that standard estimators are fundamentally biased due to this finite binning; we will address this in a future work.
+%TODO: figure out this wording.
+need to explain more than this or leave out entirely?
 
 %some papers have equations (even this eqn) in intro - but should bump it to motivation and just explain?
-The current standard estimator was proposed by \cite{LandySzalay1993}, hereafter LS, which is based on summing all data pairs $DD$ with a given separation and using data-random pairs $DR$ and random pairs $RR$ to correct for the survey boundary.
+The current standard estimator was proposed by \cite{LandySzalay1993}, hereafter \LS.
+It is based on summing all data pairs $DD$ with a given separation and using data-random pairs $DR$ and random pairs $RR$ to correct for the survey boundary.
 The correlation function $\xi_k(r)$ for the $k^\mathrm{th}$ separation bin is
@@
 \begin{equation}
-\xi_k(r) = \frac{DD_k(r) - 2DR_k(r) + RR_k(r)}{RR_k(r)}.
+\xi_k(r) = \frac{DD_k(r) - 2DR_k(r) + RR_k(r)}{RR_k(r)},
 \end{equation}
-Compared with other estimators based on combinations of $DD$, $DR$ and $RR$, LS has been shown to have the lowest bias and variance \citep{Kerscher2000}.
-%is this necessary? not sure where it fits
+where we assume the pair counts are normalized. %should we write out normalization explicitly? here or somewhere else?
+Compared with other estimators based on simple combinations of $DD$, $DR$ and $RR$, \LS has been shown to have the lowest bias and variance \citep{Kerscher2000}.
+%added simple because there was the recent paper which looked at all linear combos and optimized, think that one might have better bias/variance properties
 Estimators of the \cf must also take into account the imperfect nature of the survey area that is in the window, including the target completeness and fiber collisions; typically each galaxy pair is assigned a weight based on these.
+%is this necessary?
+not sure where it fits
+Then, pair counts are replaced by the sum of pair weights.
 
@@
 The estimators described so far have all taken probabilistic approaches; some have also taken a likelihood approach.
-\cite{BaxterRozo2013} introduced a maximum likelihood estimator for the \cf, which achieves lower variance compared to the LS estimator, enabling finer binning and requiring a smaller random catalog for the same precision.
+\cite{BaxterRozo2013} introduced a maximum likelihood estimator for the \cf, which achieves lower variance compared to the \LS estimator, enabling finer binning and requiring a smaller random catalog for the same precision.
 
-These estimators present improvements to LS, but they all still require binning in separation.
-Some require additional computational costs or layers of complexity, so the standard formulation of LS continues to be the default estimator used in analyses.
+These estimators present improvements to \LS, but they all still require binning in separation.
+Some require additional computational costs or layers of complexity, so the standard formulation of \LS continues to be the default estimator used in most analyses.
+%make sure this is true!
+talk to cosmologists who do these analyses
 
-In this paper, we present a new estimator for the correlation function, \est, which generalizes the LS estimator to produce a continuous estimation of the \cf. 
+In this paper, we present a new estimator for the correlation function, \est, which generalizes the \LS estimator to produce a continuous estimation of the \cf. 
 \Est projects the galaxy pairs onto a set of continuous basis functions and computes the best-fit linear combination of these functions.
 The basis representation can depend on the pair separation as well as other desired properties, and can also utilize the known form of the \cf.
-For top-hat basis functions, \est exactly reduces to the LS estimator. 
+For top-hat basis functions, \est exactly reduces to the \LS estimator. 
 This estimator removes the need for binning and allows for the \cf to be represented by fewer basis functions, requiring fewer mock catalogs to compute the covariance matrix.
@@
 
+%will obviously keep changing this as paper develops
 %spell out acronyms again that were defined in the abstract?
@@
 
+The pair counts for \LS and related estimators are defined explicitly as
+%TODO: fix spacing on sides of equal sign
+\begin{eqnarray}\displaystyle
+DD_k &\equiv& \sum_{n n'} i(g_k < |x_n - x_{n'}| < h_k) \\
+DR_k &\equiv& \sum_{n m} i(g_k < |x_n - x_m| < h_k) \\
+RR_k &\equiv& \sum_{m m'} i(g_k < |x_m - x_{m'}| < h_k),
+\end{eqnarray}
+where $DD_k$ is the count of data--data tracer pairs in bin $k$ (which has bin edges $g_k$ and $h_k$), $i()$ is an indicator function, $x$ is the tracer position, the $n$ and $n'$ indices index data positions, the $m$ and $m'$ indices index random catalog positions, $DR_k$ is the count of data--random pairs, and $RR_k$ is the count of random--random pairs.
+
 The Landy-Szalay estimator is justified by\ldots
@@
 Estimating clustering is closely related to least-squares fitting.
+We are trying to find the best representation of spatial data in the space of two-point radial separation.
+Recall that the linear least-squares fit to a set of data 
+\begin{equation}
+X = [A\T C\inv A]\inv [A\T C\inv Y]
+\end{equation}
+where $X$ is the vector of best-fit parameters, $A$ is a matrix with zeroth and first order terms of $x$ data, $C$ is the covariance matrix, and $Y$ is a column vector of $y$ data.
+The second bracketed term contains the observed data; the first bracketed term weights the data by the errors.
+Naively, in the case of the \cf, the observed data is the pair counts at a given separation, and the weights are provided by the pair counts of the random catalog.
+Indeed, this gives DD/RR, the so-called natural estimator (e.g. \citealt{Kerscher2000}).
 
-From these, we can infer the form of the estimator.
+By generalizing this to , we can infer the form of the estimator.
 
@@
 
-We show that it reduces to the Landy-Szalay estimator.
+We generalize the \LS estimator defined above to any set of $K$ basis functions $f$ of the pair, so we now have
+\begin{eqnarray}\displaystyle
+DD &\equiv& \sum_{n n'} f(T_n, T_{n'}) \\
+DR &\equiv& \sum_{n m} f(T_n, T_{m}) \\
+RR &\equiv& \sum_{m m'} f(T_m, T_{m'}) \\
+QQ &\equiv& \sum_{m m'} f(T_m, T_{m'}) \cdot f\T(T_m, T_{m'}),
+\end{eqnarray}
+%introduced this T notation to be the most general but it's a bit clunky (see below)
+%what do we think about K-vector notation?
+%matrix->tensor?
+where $DD$, $DR$, and $RR$ are now $K$-vectors, $T$ refers to the data for the given tracer, and $QQ$ is a $K$-by-$K$ matrix defined as the outer product of basis function evaluations of the random--random pairs.
+
+Then, we can compute the \cf as
+\begin{eqnarray}\displaystyle
+% here is where the T notation is weird - when you look at evaluation
+\xi(T_l, T_{l'}) &\equiv& a\T \cdot f(T_l, T_{l'}) \\
+a &\equiv& QQ\inv \cdot (DD + 2DR - RR) 
+\end{eqnarray}
+where $T_l$ and $T_{l'}$ contain the data values at which to evaluate $\xi$, and $a$ is a $K$-vector of the ``amplitudes'' of the basis functions.
+This generalized two-point estimator, or \est, removes the need for binning as the basis functions can be continuous.
+It is invariant under affine transformations; we show this in Appendix \ref{sec:affine}.
+
+\iffalse %This is in index notation - a lot messier and maybe inaccurate (with the inverse QQ)? but more consistent
+\begin{eqnarray}\displaystyle
+DD_k &\equiv& \sum_{n n'} f_k(T_n, T_{n'}) \\
+DR_k &\equiv& \sum_{n m} f_k(T_n, T_{m}) \\
+RR_k &\equiv& \sum_{m m'} f_k(T_m, T_{m'}) \\
+QQ_{kl} &\equiv& \sum_{m m'} f_k(T_m, T_{m'}) f_l(T_m, T_{m'}),
+\end{eqnarray}
+%should we change the letter of k to mean basis functions rather than bins now?
+%better notation for T?
+Then, we can compute the \cf as
+\begin{eqnarray}\displaystyle
+\xi_k(T_j, T_{j'}) &\equiv& a_k f_k(T_j, T_{j'}) \\
+a_k &\equiv& [QQ\inv]_{kl} (DD_l + 2DR_l - RR_l) 
+\end{eqnarray}
+where $k$ now indexes the basis functions, $T$ refers to all of the data for the given tracer, and $QQ_{kl}$ is the product of basis function indexed by $k$ and $l$ for the random--random pairs.
+\fi
+
+In the case where the basis functions only depend on the pair separation, we have $f = f(|x_n - x_{n'}|)$ and $\xi = \xi(r)$, where $r$ is the separation at which to evaluate the correlation function.
+Specifically, we can choose to define a set of basis functions $f$ as,
+\begin{equation}
+f_k(|x_n - x_n'|) =  i(g_k < |x_n - x_{n'}| < h_k)
+\end{equation}
+where $k$ denotes a particular bin in separation.
+In this case the $DD$, $DR$ and $RR$ vectors become binned pair counts and the $QQ$ matrix becomes diagonal, with diagonal elements equal to the $RR$ vector elements. %TODO CONTINUE HERE
+Then for this function $f$, which is the definition of a top-hat (rectangular) function, the estimator is equivalent to that of Landy-Szalay.
 
@@
 These show that the estimator is correct.
-
-We show that it extracts Fisher-optimal information.
 
@@
 
-We note that our estimator doesn't rely on LS; it could be implemented for various linear combination of pair counts, given certain properties. %aka the tensor term
+We note that our estimator doesn't rely on \LS; it could be implemented for various linear combination of pair counts, given certain properties. %aka the tensor term
 
@@
 
+\appendix
+\section{Affine Invariance}\label{sec:affine}
+
+The estimator is invariant under an affine transformation. We represent this by a matrix M, such that 
+\begin{equation}
+f' \leftarrow Mf
+\end{equation}
+Then in the primed basis, the pair counts become
+\begin{eqnarray}\displaystyle
+DD' &=& \sum_{n n'} M f_{n n'} = M\,DD
+\\
+DR' &=& \sum_{n m} M f_{n m} = M\,DR
+\\
+RR' &=& \sum_{m m'} M f_{m m'} = M\,RR
+\end{eqnarray}
+where $f_{n n'} = f_k(T_n, T_{n'})$.
+We have factored $M$ out of the summation. For the $QQ$ matrix we have
+\begin{eqnarray}\displaystyle
+QQ' &=& \sum_{m m'} (Mf_{m m'}) \cdot (Mf_{m m'})\T \\
+&=& M\Bigg[ \sum_{m m'} f_{m m'} \cdot f_{m m'}\T \Bigg]M\T \\
+&=& M QQ M\T
+\end{eqnarray}
+Then the amplitudes in the primed basis become
+\begin{eqnarray}\displaystyle
+a' &=& [M QQ M\T]\inv \cdot [MDD - 2\,MDR + MRR] \\
+&=& (M\T)\inv QQ\inv \cdot [DD - 2\,DR + RR] \\
+&=& (M\T)\inv a
+\end{eqnarray}
+and the estimator in the primed basis is 
+\begin{eqnarray}\displaystyle
+\xi' &=& [(M\T)\inv a]\T \cdot (Mf) \\
+&=& a\T[(M\inv)\T]\T \cdot (Mf) \\
+&=& a\T M\inv \cdot Mf \\
+&=& a\T \cdot f = \xi.
+\end{eqnarray}
+Thus after an affine transformation of the basis function, the resulting estimator is equivalent to the estimator in the original basis.
+
+We note that this requires $M$ be invertible.
+% not sure if i understand what i wrote here
+However, any two equivalent bases must be related by the inverse of a transformation matrix, so this requirement is already satisfied.
+
 %need style file error

### 2020-01-21 paper/paper.tex
paper updates

@@
 \usepackage[sort&compress]{natbib}
+\usepackage{graphicx}
+\graphicspath{ {./images/} }
 \usepackage{xspace}
+\usepackage{xcolor}
 
@@
 \newcommand{\etc}{\foreign{etc}}
-
 % math
@@
 \newcommand{\T}{^{\mathsf{T}}}
+\newcommand{\hmpc}{$h^{-1}$Mpc}
+% comments
+\newcommand{\KSF}[1]{\textcolor{teal}{KSF says: #1}}
 
@@
 \begin{document}\sloppy\sloppypar\raggedbottom\frenchspacing
-\title{\textbf{Two-point statistics without bins: A correlation function estimator for large-scale structure in a continuous-function basis}}
+
+\title{\textbf{Two-point statistics without bins: A continuous-function generalization of the correlation function estimator for large-scale structure}}
+%Two-point statistics without bins: A correlation function estimator for large-scale structure in a continuous-function basis
+%Two-point statistics without bins: The correlation function estimator for large-scale structure generalized to a continuous-function basis
 %\title{\textbf{A Continuous-Function Estimator of the Correlation Function for Large-Scale Structure}}
@@
 The two-point correlation function (\cf) is the most important statistic in structure formation.
-Current estimators of the \cf, used to measure the clustering of density field tracers, have significant limitations.
+Current estimators of the \cf, which is used to measure the clustering of density field tracers, have significant limitations.
 The standard Landy-Szalay (\LS) estimator evaluates the \cf in bins of radial separation between objects, and the choice of bins introduces a trade-off between bias and variance.
+%is radial the right word?
+does that imply line-of-sight?
 % Aims
@@
 The clustering of these objects is measured with two-point statistics, namely the power spectrum $P(k)$ and the two-point correlation function (\cf). %name xi_r?
-These characterize the clustering in Fourier space and real space, respectively, with the \cf defined as the Fourier Transform of the power spectrum; in principle, they contain the same information.
-%show eqn?
+These characterize the clustering in Fourier space and real space, respectively, with the \cf defined as the Fourier Transform of the power spectrum: 
+%TODO: CHECK, PROBS WRONG
+\begin{equation}
+\xi(r) = \frac{1}{2\pi} \int_0^{\infty} P(k) d^3k
+\end{equation}
+In principle, the power spectrum and the two-point correlation function contain the same information.
 However, in practical applications the survey boundaries introduce nontrivial issues in computing these statistics, leading to diverging approaches to their computation with a significant difference in expense.
@@
 Naively, in the case of the \cf, the observed data is the pair counts at a given separation, and the weights are provided by the pair counts of the random catalog.
-Indeed, this gives DD/RR, the so-called natural estimator (e.g. \citealt{Kerscher2000}).
+Indeed, this is reminiscent of the so-called natural estimator of the \cf, $\xi_k = DD_k/RR_k - 1$ (e.g. \citealt{Kerscher2000}).
 
-By generalizing this to , we can infer the form of the estimator.
+From this connection, we can infer the form of the estimator.
 
 % TODO: fix this capitalization issue
-\section{\Est} \label{sec:est}
-
-We derive/define this new estimator.
+\section{\Est} 
+\label{sec:est}
 
@@
 
-\iffalse %This is in index notation - a lot messier and maybe inaccurate (with the inverse QQ)? but more consistent
+%%% INDEX NOTATION VERSION %%%
+\iffalse % Messier and maybe inaccurate (with the inverse QQ)? but more consistent
 \begin{eqnarray}\displaystyle
@@
 where $k$ now indexes the basis functions, $T$ refers to all of the data for the given tracer, and $QQ_{kl}$ is the product of basis function indexed by $k$ and $l$ for the random--random pairs.
-\fi
+\fi %%%%%%%%%%%%%%%%%%
 
@@
 
-We show its performance on a simulated Gaussian random field.
+\subsection{Demonstration using Spline Basis Functions}
+\label{sec:spline}
 
-\section{Application to survey data} \label{sec:app}
+We demonstrate the application of the \est on a simulated Gaussian random field.
+We generate lognormal mock catalogs using the \texttt{nbodykit} package.
+We use an input power spectrum with the Planck cosmology \KSF{CITE}.
+The true correlation function is then known: it is the Fourier transform of the input power spectrum, computed numerically.
+Our test catalogs have size (750 \hmpc)$^3$ and a galaxy number density of $3 \times 10^{-4}$, to match that of the Brightest Cluster Galaxies \KSF{CITE}. 
+We compute 100 realizations of this box. 
+We also generate a uniformly sampled random catalog with 10 times the number of galaxies as in the data catalogs.
 
-We apply the estimator to the SDSS DR3 sample from \cite{Eisenstein2005}.
+\label{fig:quadratic}
+\begin{figure}[h]
+\centering
+    \includegraphics[width=0.8\textwidth]{tophat_quadratic}
+    \caption{A comparison between the standard tophat estimator (blue solid) and the \est with a quadratic spline basis function (red dotted).
+The shaded region is the 1$\sigma$ variation in the 100 mock catalogs.
+The quadratic estimator has lower bias and lower variance with fewer bins.}
+\end{figure}
 
-The reconstructed dataset we use is described in \cite{Kazin2010}.
+\KSF{Should i have a figure showing the basis functions?}
 
-We use \est to linearize around the BAO feature.
+\KSF{These sims have the weird thing where the data says the peak is offset. Should figure out, and also it's distracting from the main point of the figure}
+
+\KSF{FINAL PARAMS:  number density 3e-4 (now 1e-4). rmax 200 hMpc (now 150). rcont=1000 (now rcont=300). L=1000 hMpc (now 750 - need more to account for higher rmax)}
+
+\KSF{should RMSE be shown in figure as well as described in text?}
+
+We compare the standard estimator, reformulated as continuous functions using a tophat basis, to the \est with a quadratic spline basis.
+We use second-order B-splines, which constitute the set of basis functions for a quadratic spline \KSF{CITE}.
+For the tophat basis we use 44 basis functions in the range $40 < r < 150$ \hmpc, each with a width of 2.5 \hmpc. 
+For the quadratic spline basis, we use the same $r$ range, but with only 11 basis functions, and knots chosen on a grid of 10 \hmpc.
+\KSF{need to specify more about knots? is this accurate (not quite, the knots repeat values at the edge i think...) and should i mention control points?}
+
+The results are shown in Figure~\ref{fig:quadratic}, compared to the true input \cf.
+The quadratic basis results in a \cf estimate that has a root mean square error (RMSE) with respect to the truth of $4.39 \times 10^{-4}$, compared to $5.43 \times 10^{-4}$ for the tophat basis.
+This holds true when we average over the bin \KSF{need to do weighted average based on which r-values contribute? supposed to hear from tinker}, as in standard practice, and then compute the RMSE.
+The quadratic basis also results in lower variance across the 100 mocks, compared to the tophat basis.
+Thus, the \est allows for a more accurate and precise estimate of the \cf with fewer components.
+
+\subsection{BAO Scale Estimation Test}
+
+Measurement of the BAO scale provides a good use case for our estimator.
+
+The BAO peak position is a function of multiple distance scales.
+We can measure it by comparing the data to a fiducial model of clustering.
+To a very good approximation, the peak position depends on the single scale dilation parameter
+\begin{equation}
+\alpha = \Bigg( \frac{D_A(z)}{D_A^{mod}(z)} \Bigg)^{2/3} \Bigg( \frac{H^{mod}(z)}{H(z)} \Bigg)^{1/3} \Bigg( \frac{r_s^{mod}}{r_s} \Bigg),
+\end{equation}
+where $D_A$ is the angular diameter distance, $H$ is the Hubble constant, $r_s$ is the sound horizon scale at the drag epoch, and the superscript ``$mod$'' denotes the value for the fiducial model.
+If we find $\alpha>1$, the peak is shifted to smaller scales than its position in the model, and if $\alpha<1$, the peak is shifted to larger scales.
+
+For BAO measurements, we measure the spherically averaged redshift space correlation function, $\xi(s)$.
+This multipoles of the correlation function are given by
+\begin{equation}
+\xi_l(s) = \frac{1}{2} \int^{1}_{-1} \xi_l(s, \mu) L_l(\mu) d\mu
+\end{equation}
+where $L_l$ is the Legendre multipole of order $l$,  $s$ is the redshift-space separation between pairs, and $\mu = cos(\theta)$, $\theta$ is the angle between the separation vector $s$ and the line-of-sight direction (i.e.
+$\mu=1$ indicates a galaxy pair along the line of sight, $\mu$=0 indicates a pair with only transverse separation).
+The two-point correlation function that we use is the $l=0$ multiple; from now on we will omit the subscript, $\xi(s) = \xi_0(s)$.
+
+In standard practice, a fitting function is used to determine the value of $\alpha$:
+\begin{equation}
+\xi^{fit}(s) = B^2 \xi^{mod}(\alpha s) + \frac{a_1}{s^2} + \frac{a_2}{s} + a_3
+\end{equation}
+where $B$ is a constant that allows for a large-scale bias, and $a_1$, $a_2$, and $a_3$ are nuisance parameters to account for the broadband shape.
+The model is given by \KSF{DESCRIBE}, and a $\chi^2$ fit is performed with five free parameters: $\alpha$, $B$, $a_1$, $a_2$, and $a_3$.
+The resulting value for $\alpha$ is used to derive the actual values of the distance scales of interest.
+
+The form of the standard fitting function is well-suited to our estimator, as it is a few-parameter model with a linear combination of terms.
+To use our estimator to estimate $\alpha$, we take the numerical partial derivative of the model with respect to $\alpha$, using a change in $\alpha$ of size $d\alpha$.
+Our fitting function is then
+\KSF{Check if the derivative part is right!}
+\begin{equation}
+\xi^{fit}(s) = B^2 \xi^{mod}(s) + C d\alpha\frac{d\xi_{mod}(\alpha s)}{d\alpha} + \frac{a_1}{s^2} + \frac{a_2}{s} + a_3,
+\end{equation}
+where $C$ describes the contribution of $\alpha$.
+That is, a value of $C=0$ indicates that the model is the best fit to the data and no scale dilation is needed.
+More generally, we can use $C$ to compute $\alpha$ with $\alpha = 1 + Cd\alpha$.
+With the terms in this fitting function the bases of our estimator, the values $B$, $C$, $a_1$, $a_2$, and $a_3$ are then given by the output amplitude vector $a$ as described in $\S$\ref{sec:est}.
+
+\KSF{Choose realistic alpha values}
+
+\KSF{Instead of this alpha test, should do a slightly more realistic one where we change the model cosmology? Issue is that we then don't know the truth}
+
+\KSF{BUT my estimator is having a hard time recovering the truth anyway!
+Because of variance in the realizations.
+It converges to the same, but wrong, value of alpha, and diff ones for diff realizations.
+Need to average over many?
+(Still maybe wrong due to some persistant weirdness as mentioned above!
+Maybe need to check how I'm doing the lognormals)}
+
+\label{fig:bao_bases}
+\begin{figure}[h]
+\centering
+    \includegraphics[width=0.8\textwidth]{bao_bases}
+    \caption{The set of basis functions used to fit for the BAO scale using our estimator.
+The $B^2$ term (green) is the fiducial model used to determine the scale dilation parameter $\alpha$.
+The $C$ term is the derivative of this model with respect to $\alpha$, allowing for the estimation of this parameter.
+The $a_1$, $a_2$, and $a_3$ terms are nuisance parameters to fit the broadband shape.
+The \KSF{Make colorblind friendly!
+Do i also need to use different linestyles?
+Or maybe weights?}}
+\end{figure}
+
+We demonstrate this method using the same set of lognormal mocks as in $\S$\ref{sec:spline}.
+We construct a simple test in which our model is the known input model, the Planck parameters, but with some scale dilation offset, here $\alpha_{mod}=1.05$ (while the true value is, of course, $\alpha=1$).
+The set of basis functions for our estimator is shown in Figure~\ref{fig:bao_bases}.
+
+We perform an iterative procedure to estimate $\alpha$.
+After the first estimation with the $\alpha=1.05$, model, we use the value of $C$ to compute the recovered $\alpha$, as described above.
+We then take that value as our $\alpha_{mod}$, and re-run the estimator.
+We stop when the percent change between $\alpha_{mod}$ and the recovered $\alpha$ is less than 0.01\%; it typically takes only $\sim$5 iterations to converge.
+The resulting estimate for the correlation function is shown in Figure~\ref{fig:bao}, in comparison to the traditional tophat estimator.
+
+\label{fig:bao}
+\begin{figure}[h]
+\centering
+    \includegraphics[width=0.8\textwidth]{bao}
+    \caption{Estimation of the correlation function using our estimator with basis functions based on the BAO fitting function (purple dot-dashed).
+We also show the standard Landy-Szalay estimator, displayed as a tophat function (blue), as well as the true input correlation function (black).
+\KSF{This is a placeholder figure!
+Need to resolve issues above.} \KSF{Should this and the basis set figure be a single figure with two panels?}}
+\end{figure}
+
+% \section{Application to survey data} \label{sec:app}
+
+% We apply the estimator to the SDSS DR3 sample from \cite{Eisenstein2005}.
+
+% The reconstructed dataset we use is described in \cite{Kazin2010}.
 
@@
 
+\subsection{Beyond the Landy-Szalay Estimator}
+
 We note that our estimator doesn't rely on \LS; it could be implemented for various linear combination of pair counts, given certain properties. %aka the tensor term
+
+We discuss the relation of our estimator to the optimized estimator of \cite{VargasMagana2013}, which also uses a linear combination approach.
+
+\subsection{Computational Performance}
 
@@
 
-We discuss the impact on computing the covariance matrix.
+\subsection{Computing the Covariance Matrix}
 
-We discuss the relation of our estimator to the optimized estimator of \cite{VargasMagana2013}, which also uses a linear combination approach.
+The \est results in \cf estimates that are just as accurate with fewer components.
+This is critical when computing the covariance matrix.
+The covariance matrix is used to estimate the error on a measurement, by evaluating the \cf on a large number of mock catalogs and computing the covariance between the bins.
+In our replacement of bins with continuous functions, the covariance matrix will be the covariance between these functions.
+The unbiased estimator for the sample covariance matrix is
+\begin{equation}
+\hat{C}^{ML}_{ij} = \frac{1}{N_{mocks}-1} \sum_{k=1}^{N_{mocks}} \bigg( x^k_i - \bar{x}_i \bigg) \bigg( x^k_j - \bar{x}_j \bigg),
+\end{equation}
+where $k$ denotes the index of the mock, $i$ and $j$ denote the index of the bin or component, $x$ denotes the estimate in that bin for that mock, and $\bar{x}$ denotes the mean value of the estimate in that bin across the mocks.
+To get an unbiased estimate of the inverse covariance matrix, we require a correction factor, as the inverse of an unbiased estimator is not necessarily unbiased.
+The unbiased estimator for the sample inverse covariance matrix can be shown to be \citep{Hartlap2007}
+\begin{equation}
+\hat{C}\inv = \frac{N_{mocks}-N_{bins}-2}{N_{mocks}-1} \bigg( \hat{C}^{ML}\bigg) \inv.
+\end{equation}
+
+%This prefactor correction results in the  propagates to the variance in 
+The variance in the elements of this estimator then have a dependence on $N_{mocks}$ and $N_{bins}$.
+This propagates to the derived cosmological parameters, resulting in an overestimation of the error bars.
+Assuming that $N_{mocks} >> N_{bins}$ (and both much larger than the number of parameters to be estimated), and that the measurements are Gaussian distribued, the error bars are inflated by a factor of $(1 + N_{bins}/N_{mocks})$ (i.e., the true constraints are tighter than the derived ones).
+This factor is critical at the precision of cosmological parameter estimation.
+
+Typically, this is dealt with by generating a very large number of mocks.
+For the Baryon Oscillation Spectroscopic Survey (BOSS, \citealt{Eisenstein2011}), $\sim$600 mocks were needed and the analysis used 41 bins.
+Future surveys will have more costly requirements on mock catalogs, with larger simulations necessary to cover the larger survey volumes.
+
+An alternative to increasing $N_{mocks}$ is decreasing $N_{bins}$ to achieve the same error on precision.
+In the standard method, this is shown to \emph{increase} the statistical error, albeit only slightly \cite{Percival2013}.
+A substantial increase in bin width would prevent capturing information in finer clustering features; even the relatively broad BAO peak requires a bin size on the order of its width of $\sim$10\hmpc.
+In fact, in the standard method more bins would be desireable, but the number is limited by the available number of mocks due to the error discussed above.
+
+With our estimator, we have shown that we can reduce the variance by using fewer components, without sacrificing accuracy.
+This means that we can safely reduce $N_{bins}$.
+To then achieve the same precision on the error on the cosmological parameters, a lower value of $N_{mocks}$ becomes possible.
+This will significantly reduce requirements on mocks, which will be particularly important for upcoming large surveys.
+
+\subsection{Further Applications}
 

### 2020-01-21 paper/sample62.tex
paper updates

%%
%% Beginning of file 'sample62.tex'
%%
%% Modified 2018 January
%%
%% This is a sample manuscript marked up using the
%% AASTeX v6.2 LaTeX 2e macros.
%%
%% AASTeX is now based on Alexey Vikhlinin's emulateapj.cls 
%% (Copyright 2000-2015).  See the classfile for details.

%% AASTeX requires revtex4-1.cls (http://publish.aps.org/revtex4/) and
%% other external packages (latexsym, graphicx, amssymb, longtable, and epsf).
%% All of these external packages should already be present in the modern TeX 
%% distributions.  If not they can also be obtained at www.ctan.org.

%% The first piece of markup in an AASTeX v6.x document is the \documentclass
%% command. LaTeX will ignore any data that comes before this command. The 
%% documentclass can take an optional argument to modify the output style.
%% The command below calls the preprint style  which will produce a tightly 
%% typeset, one-column, single-spaced document.  It is the default and thus
%% does not need to be explicitly stated.
%%
%%
%% using aastex version 6.2
\documentclass{aastex62}

%% The default is a single spaced, 10 point font, single spaced article.
%% There are 5 other style options available via an optional argument. They
%% can be envoked like this:
%%
%% \documentclass[argument]{aastex62}
%% 
%% where the layout options are:
%%
%%  twocolumn   : two text columns, 10 point font, single spaced article.
%%                This is the most compact and represent the final published
%%                derived PDF copy of the accepted manuscript from the publisher
%%  manuscript  : one text column, 12 point font, double spaced article.
%%  preprint    : one text column, 12 point font, single spaced article.  
%%  preprint2   : two text columns, 12 point font, single spaced article.
%%  modern      : a stylish, single text column, 12 point font, article with
%% 		  wider left and right margins. This uses the Daniel
%% 		  Foreman-Mackey and David Hogg design.
%%  RNAAS       : Preferred style for Research Notes which are by design 
%%                lacking an abstract and brief. DO NOT use \begin{abstract}
%%                and \end{abstract} with this style.
%%
%% Note that you can submit to the AAS Journals in any of these 6 styles.
%%
%% There are other optional arguments one can envoke to allow other stylistic
%% actions. The available options are:
%%
%%  astrosymb    : Loads Astrosymb font and define \astrocommands. 
%%  tighten      : Makes baselineskip slightly smaller, only works with 
%%                 the twocolumn substyle.
%%  times        : uses times font instead of the default
%%  linenumbers  : turn on lineno package.
%%  trackchanges : required to see the revision mark up and print its output
%%  longauthor   : Do not use the more compressed footnote style (default) for 
%%                 the author/collaboration/affiliations. Instead print all
%%                 affiliation information after each name. Creates a much
%%                 long author list but may be desirable for short author papers
%%
%% these can be used in any combination, e.g.
%%
%% \documentclass[twocolumn,linenumbers,trackchanges]{aastex62}
%%
%% AASTeX v6.* now includes \hyperref support. While we have built in specific
%% defaults into the classfile you can manually override them with the
%% \hypersetup command. For example,
%%
%%\hypersetup{linkcolor=red,citecolor=green,filecolor=cyan,urlcolor=magenta}
%%
%% will change the color of the internal links to red, the links to the
%% bibliography to green, the file links to cyan, and the external links to
%% magenta. Additional information on \hyperref options can be found here:
%% https://www.tug.org/applications/hyperref/manual.html#x1-40003
%%
%% If you want to create your own macros, you can do so
%% using \newcommand. Your macros should appear before
%% the \begin{document} command.
%%
\newcommand{\vdag}{(v)^\dagger}
\newcommand\aastex{AAS\TeX}
\newcommand\latex{La\TeX}

%% Reintroduced the \received and \accepted commands from AASTeX v5.2
\received{January 1, 2018}
\revised{January 7, 2018}
\accepted{\today}
%% Command to document which AAS Journal the manuscript was submitted to.
%% Adds "Submitted to " the arguement.
\submitjournal{ApJ}

%% Mark up commands to limit the number of authors on the front page.
%% Note that in AASTeX v6.2 a \collaboration call (see below) counts as
%% an author in this case.
%
%\AuthorCollaborationLimit=3
%
%% Will only show Schwarz, Muench and "the AAS Journals Data Scientist 
%% collaboration" on the front page of this example manuscript.
%%
%% Note that all of the author will be shown in the published article.
%% This feature is meant to be used prior to acceptance to make the
%% front end of a long author article more manageable. Please do not use
%% this functionality for manuscripts with less than 20 authors. Conversely,
%% please do use this when the number of authors exceeds 40.
%%
%% Use \allauthors at the manuscript end to show the full author list.
%% This command should only be used with \AuthorCollaborationLimit is used.

%% The following command can be used to set the latex table counters.  It
%% is needed in this document because it uses a mix of latex tabular and
%% AASTeX deluxetables.  In general it should not be needed.
%\setcounter{table}{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% The following section outlines numerous optional output that
%% can be displayed in the front matter or as running meta-data.
%%
%% If you wish, you may supply running head information, although
%% this information may be modified by the editorial offices.
\shorttitle{Sample article}
\shortauthors{Schwarz et al.}
%%
%% You can add a light gray and diagonal water-mark to the first page 
%% with this command:
% \watermark{text}
%% where "text", e.g. DRAFT, is the text to appear.  If the text is 
%% long you can control the water-mark size with:
%  \setwatermarkfontsize{dimension}
%% where dimension is any recognized LaTeX dimension, e.g. pt, in, etc.
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% This is the end of the preamble.  Indicate the beginning of the
%% manuscript itself with \begin{document}.

\begin{document}

\title{An Example Article using \aastex v6.2\footnote{Released on January, 8th, 2018}}

%% LaTeX will automatically break titles if they run longer than
%% one line. However, you may use \\ to force a line break if
%% you desire. In v6.2 you can include a footnote in the title.

%% A significant change from earlier AASTEX versions is in the structure for 
%% calling author and affilations. The change was necessary to implement 
%% autoindexing of affilations which prior was a manual process that could 
%% easily be tedious in large author manuscripts.
%%
%% The \author command is the same as before except it now takes an optional
%% arguement which is the 16 digit ORCID. The syntax is:
%% \author[xxxx-xxxx-xxxx-xxxx]{Author Name}
%%
%% This will hyperlink the author name to the author's ORCID page. Note that
%% during compilation, LaTeX will do some limited checking of the format of
%% the ID to make sure it is valid.
%%
%% Use \affiliation for affiliation information. The old \affil is now aliased
%% to \affiliation. AASTeX v6.2 will automatically index these in the header.
%% When a duplicate is found its index will be the same as its previous entry.
%%
%% Note that \altaffilmark and \altaffiltext have been removed and thus 
%% can not be used to document secondary affiliations. If they are used latex
%% will issue a specific error message and quit. Please use multiple 
%% \affiliation calls for to document more than one affiliation.
%%
%% The new \altaffiliation can be used to indicate some secondary information
%% such as fellowships. This command produces a non-numeric footnote that is
%% set away from the numeric \affiliation footnotes.  NOTE that if an
%% \altaffiliation command is used it must come BEFORE the \affiliation call,
%% right after the \author command, in order to place the footnotes in
%% the proper location.
%%
%% Use \email to set provide email addresses. Each \email will appear on its
%% own line so you can put multiple email address in one \email call. A new
%% \correspondingauthor command is available in V6.2 to identify the
%% corresponding author of the manuscript. It is the author's responsibility
%% to make sure this name is also in the author list.
%%
%% While authors can be grouped inside the same \author and \affiliation
%% commands it is better to have a single author for each. This allows for
%% one to exploit all the new benefits and should make book-keeping easier.
%%
%% If done correctly the peer review system will be able to
%% automatically put the author and affiliation information from the manuscript
%% and save the corresponding author the trouble of entering it by hand.

\correspondingauthor{August Muench}
\email{greg.schwarz@aas.org, gus.muench@aas.org}

\author[0000-0002-0786-7307]{Greg J. Schwarz}
\affil{American Astronomical Society \\
2000 Florida Ave., NW, Suite 300 \\
Washington, DC 20009-1231, USA}

\author{August Muench}
\affiliation{American Astronomical Society \\
2000 Florida Ave., NW, Suite 300 \\
Washington, DC 20009-1231, USA}
\collaboration{(AAS Journals Data Scientists collaboration)}

\author{Butler Burton}
\affiliation{National Radio Astronomy Observatory}
\affiliation{AAS Journals Associate Editor-in-Chief}
\nocollaboration

\author{Amy Hendrickson}
\altaffiliation{Creator of AASTeX v6.2}
\affiliation{TeXnology Inc.}
\collaboration{(LaTeX collaboration)}

\author{Julie Steffen}
\affiliation{AAS Director of Publishing}
\affiliation{American Astronomical Society \\
2000 Florida Ave., NW, Suite 300 \\
Washington, DC 20009-1231, USA}

\author{Jeff Lewandowski}
\affiliation{IOP Senior Publisher for the AAS Journals}
\affiliation{IOP Publishing, Washington, DC 20005}

%% Note that the \and command from previous versions of AASTeX is now
%% depreciated in this version as it is no longer necessary. AASTeX 
%% automatically takes care of all commas and "and"s between authors names.

%% AASTeX 6.2 has the new \collaboration and \nocollaboration commands to
%% provide the collaboration status of a group of authors. These commands 
%% can be used either before or after the list of corresponding authors. The
%% argument for \collaboration is the collaboration identifier. Authors are
%% encouraged to surround collaboration identifiers with ()s. The 
%% \nocollaboration command takes no argument and exists to indicate that
%% the nearby authors are not part of surrounding collaborations.

%% Mark off the abstract in the ``abstract'' environment. 
\begin{abstract}

This example manuscript is intended to serve as a tutorial and template for
authors to use when writing their own AAS Journal articles. The manuscript
includes a history of \aastex\ and documents the new features in the
previous versions as well as the new features in version 6.2. This
manuscript includes many figure and table examples to illustrate these new
features.  Information on features not explicitly mentioned in the article
can be viewed in the manuscript comments or more extensive online
documentation. Authors are welcome replace the text, tables, figures, and
bibliography with their own and submit the resulting manuscript to the AAS
Journals peer review system.  The first lesson in the tutorial is to remind
authors that the AAS Journals, the Astrophysical Journal (ApJ), the
Astrophysical Journal Letters (ApJL), and Astronomical Journal (AJ), all
have a 250 word limit for the abstract\footnote{Note that manuscripts 
submitted to the new Research Notes of the American Astronomical Society 
(RNAAS) do \textbf{not} have abstracts.}.  If you exceed this length the
Editorial office will ask you to shorten it.

\end{abstract}

%% Keywords should appear after the \end{abstract} command. 
%% See the online documentation for the full list of available subject
%% keywords and the rules for their use.
\keywords{editorials, notices --- 
miscellaneous --- catalogs --- surveys}

%% From the front matter, we move on to the body of the paper.
%% Sections are demarcated by \section and \subsection, respectively.
%% Observe the use of the LaTeX \label
%% command after the \subsection to give a symbolic KEY to the
%% subsection for cross-referencing in a \ref command.
%% You can use LaTeX's \ref and \label commands to keep track of
%% cross-references to sections, equations, tables, and figures.
%% That way, if you change the order of any elements, LaTeX will
%% automatically renumber them.
%%
%% We recommend that authors also use the natbib \citep
%% and \citet commands to identify citations.  The citations are
%% tied to the reference list via symbolic KEYs. The KEY corresponds
%% to the KEY in the \bibitem in the reference list below.

\section{Introduction} \label{sec:intro}

\latex\ \footnote{\url{http://www.latex-project.org/}} is a document markup
language that is particularly well suited for the publication of
mathematical and scientific articles \citep{lamport94}. \latex\ was written
in 1985 by Leslie Lamport who based it on the \TeX\ typesetting language
which itself was created by Donald E. Knuth in 1978.  In 1988 a suite of
\latex\ macros were developed to investigate electronic submission and
publication of AAS Journal articles \citep{1989BAAS...21..780H}.  Shortly
afterwards, Chris Biemesdefer merged these macros and more into a \latex\
2.08 style file called \aastex.  These early \aastex\ versions introduced
many common commands and practices that authors take for granted today.
Substantial revisions
were made by Lee Brotzman and Pierre Landau when the package was updated to
v4.0.  AASTeX v5.0, written in 1995 by Arthur Ogawa, upgraded to \latex\ 2e
which uses the document class in lieu of a style file.  Other improvements
to version 5 included hypertext support, landscape deluxetables and
improved figure support to facilitate electronic submission.  
\aastex\ v5.2 was released in 2005 and introduced additional graphics
support plus new mark up to identifier astronomical objects, datasets and
facilities.

In 1996 Maxim Markevitch modified the AAS preprint style file, aaspp4.sty,
to closely emulate the very tight, two column style of a typeset
Astrophysical Journal article.  The result was emulateapj.sty.  A year
later Alexey Vikhlinin took over development and maintenance.  In 2001 he
converted emulateapj into a class file in \latex\ 2e and in 2003 Vikhlinin
completely rewrote emulateapj based on the APS Journal's RevTEX class.

During this time emulateapj gained growing acceptance in the astronomical
community as it filled an author need to obtain an approximate number of
manuscript pages prior to submission for cost and length estimates. The
tighter typeset also had the added advantage of saving paper when printing
out hard copies.

Even though author publication charges are no longer based on print pages
\footnote{see Section \ref{sec:pubcharge} in the Appendix for more details
about how current article costs are calculated.} the emulateapj class file
has proven to be extremely popular with AAS Journal authors.  An informal
analysis of submitted \latex\ manuscripts in 2015 revealed that $\sim$65\%
either called emulateapj or have a commented emulateapj classfile call
indicating it was used at some stage of the manuscript construction.
Clearly authors want to have access to a tightly typeset version of the
article when corresponding with co-authors and for preprint submissions.

When planning the next \aastex\ release the popularity of emulateapj played
an important roll in the decision to drop the old base code and adopt and
modify emulateapj for \aastex\ v6.+ instead.  The change brings \aastex\
inline with what the majority of authors are already using while still
delivering new and improved features.  \aastex\ v6.0 through v6.2 were written
by Amy Hendrickson and released in January 2016 (v6.0), October 2016 (v6.1),
and January 2018 (v6.2), respectively.  Some of the new features in v6.0 
included:

\begin{enumerate}
\item improved citations for third party data repositories and software,
\item easier construction of matrix figures consisting of multiple 
encapsulated postscript (EPS) or portable document format (PDF) files,
\item figure set mark up for large collections of similar figures,
\item color mark up to easily enable/disable revised text highlighting,
\item improved url support, and
\item numerous table options such as the ability to hide columns, column 
decimal alignment, automatic column math mode and numbering, plus splitting of
wide tables.
\end{enumerate}

The features in v6.1 were:

\begin{enumerate}
\item ORCID support for preprints,
\item improved author, affiliation and collaboration mark up,
\item reintroduced the old AASTeX v5.2 {\tt\string\received}, 
      {\tt\string\revised}, {\tt\string\accepted}, and
      {\tt\string\published} commands plus
      added the new {\tt\string\submitjournal} command to document
      which AAS Journal the manuscript was submitted to, plus
\item new typeset style options.
\end{enumerate}

The new features in v6.2 are:

\begin{enumerate}
\item A new RNAAS style option for Research Note manuscripts,
\item Titles no longer put in all caps,
\item No page skip between the title page and article body,
\item re-introduce RevTeX's widetext environment for long lines in
      two column style formats, and
\item upgrade to the {\tt\string\doi} command.
\end{enumerate}

The rest of this article provides information and examples on how to create
your own AAS Journal manuscript with v6.2.  Special emphasis is placed on
how to use the full potential of \aastex\ v6+.  The next section describes
the different manuscript styles available and how they differ from past
releases.  Section \ref{sec:floats} describes how tables and figures are
placed in a \latex\ document. Specific examples of tables, Section
\ref{subsec:tables}, and figures, Section \ref{subsec:figures}, are also
provided.  Section \ref{sec:displaymath} discusses how to display math and
incorporate equations in a manuscript while Section \ref{sec:highlight}
discuss how to use the new revision mark up.  The last section,
\ref{sec:cite}, shows how recognize software and external data as first
class references in the manuscript bibliography.  An appendix is included
to show how to construct one and provide some information on how article
charges are calculated.  Additional information is available both embedded
in the comments of this \latex\ file and in the online documentation at
\url{http://journals.aas.org/authors/aastex.html}.

\section{Manuscript styles} \label{sec:style}

The default style in \aastex\ v6.2 is a tight single column style, e.g.  10
point font, single spaced.  The single column style is very useful for
article with wide equations. It is also the easiest to style to work with
since figures and tables, see Section \ref{sec:floats}, will span the
entire page, reducing the need for address float sizing.

To invoke a two column style similar to the what is produced in
the published PDF copy use \\

\noindent {\tt\string\documentclass[twocolumn]\{aastex62\}}. \\

\noindent Note that in the two column style figures and tables will only
span one column unless specifically ordered across both with the ``*'' flag,
e.g. \\

\noindent{\tt\string\begin\{figure*\}} ... {\tt\string\end\{figure*\}}, \\
\noindent{\tt\string\begin\{table*\}} ... {\tt\string\end\{table*\}}, and \\
\noindent{\tt\string\begin\{deluxetable*\}} ... {\tt\string\end\{deluxetable*\}}. \\

\noindent This option is ignored in the onecolumn style.

Some other style options are outlined in the commented sections of this 
article.  Any combination of style options can be used.

Two style options that are needed to fully use the new revision tracking
feature, see Section \ref{sec:highlight}, are {\tt\string linenumbers} which 
uses the lineno style file to number each article line in the left margin and 
{\tt\string trackchanges} which controls the revision and commenting highlight
output.

There is also a new {\tt\string modern} option that uses a Daniel
Foreman-Mackey and David Hogg design to produce stylish, single column
output that has wider left and right margins. It is designed to have fewer
words per line to improve reader retention. It also looks better on devices
with smaller displays such as smart phones.

For a Research Note use the {\tt\string RNAAS} option which will produce a
manuscript with no abstract and in the {\tt\string modern} style.

\section{Floats} \label{sec:floats}

Floats are non-text items that generally can not be split over a page.
They also have captions and can be numbered for reference.  Primarily these
are figures and tables but authors can define their own. \latex\ tries to
place a float where indicated in the manuscript but will move it later if
there is not enough room at that location, hence the term ``float''.

Authors are encouraged to embed their tables and figures within the text as
they are mentioned.  Please do not place the figures and text at the end of
the article as was the old practice.  Editors and the vast majority of
referees find it much easier to read a manuscript with embedded figures and
tables.

Depending on the number of floats and the particular amount of text and
equations present in a manuscript the ultimate location of any specific
float can be hard to predict prior to compilation. It is recommended that
authors textbf{not} spend significant time trying to get float placement
perfect for peer review.  The AAS Journal's publisher has sophisticated
typesetting software that will produce the optimal layout during
production.

Note that authors of Research Notes are only allowed one float, either one
table or one figure.

\startlongtable
\begin{deluxetable}{c|cc}
\tablecaption{ApJ costs from 1991 to 2013\tablenotemark{a} \label{tab:table}}
\tablehead{
\colhead{Year} & \colhead{Subscription} & \colhead{Publication} \\
\colhead{} & \colhead{cost} & \colhead{charges\tablenotemark{b}}\\
\colhead{} & \colhead{(\$)} & \colhead{(\$/page)}
}
\colnumbers
\startdata
1991 & 600 & 100 \\
1992 & 650 & 105 \\
1993 & 550 & 103 \\
1994 & 450 & 110 \\
1995 & 410 & 112 \\
1996 & 400 & 114 \\
1997 & 525 & 115 \\
1998 & 590 & 116 \\
1999 & 575 & 115 \\
2000 & 450 & 103 \\
2001 & 490 &  90 \\
2002 & 500 &  88 \\
2003 & 450 &  90 \\
2004 & 460 &  88 \\
2005 & 440 &  79 \\
2006 & 350 &  77 \\
2007 & 325 &  70 \\
2008 & 320 &  65 \\
2009 & 190 &  68 \\
2010 & 280 &  70 \\
2011 & 275 &  68 \\
2012 & 150 &  56 \\
2013 & 140 &  55 \\
\enddata
\tablenotetext{a}{Adjusted for inflation}
\tablenotetext{b}{Accounts for the change from page charges to digital quanta in April, 2011}
\tablecomments{Note that {\tt \string \colnumbers} does not work with the 
vertical line alignment token. If you want vertical lines in the headers you
can not use this command at this time.}
\end{deluxetable}

For authors that do want to take the time to optimize the locations of
their floats there are some techniques that can be used.  The simplest
solution is to placing a float earlier in the text to get the position
right but this option will break down if the manuscript is altered, see
Table \ref{tab:table}.  A better method is to force \latex\ to place a
float in a general area with the use of the optional {\tt\string [placement
specifier]} parameter for figures and tables. This parameter goes after
{\tt\string \begin\{figure\}}, {\tt\string \begin\{table\}}, and
{\tt\string \begin\{deluxetable\}}.  The main arguments the specifier takes
are ``h'', ``t'', ``b'', and ``!''.  These tell \latex\ to place the float
\underline{h}ere (or as close as possible to this location as possible), at
the \underline{t}op of the page, and at the \underline{b}ottom of the page.
The last argument, ``!'', tells \latex\ to override its internal method of
calculating the float position.  A sequence of rules can be created by
using multiple arguments.  For example, {\tt\string \begin\{figure\}[htb!]}
tells \latex\ to try the current location first, then the top of the page
and finally the bottom of the page without regard to what it thinks the
proper position should be.  Many of the tables and figures in this article
use a placement specifier to set their positions.

Note that the \latex\ {\tt\string tabular} environment is not a float.  Only
when a {\tt\string tabular} is surrounded by {\tt\string\begin\{table\}} ...
{\tt\string\end\{table\}} is it a true float and the rules and suggestions
above apply.

In AASTeX v6.2 all deluxetables are float tables and thus if they are
longer than a page will spill off the bottom. Long deluxetables should
begin with the {\tt\string\startlongtable} command. This initiates a
longtable environment.  Authors might have to use {\tt\string\clearpage} to
isolate a long table or optimally place it within the surrounding text.

\begin{deluxetable*}{ccCrlc}[b!]
\tablecaption{Column math mode in an observation log \label{tab:mathmode}}
\tablecolumns{6}
\tablenum{2}
\tablewidth{0pt}
\tablehead{
\colhead{UT start time\tablenotemark{a}} &
\colhead{MJD start time\tablenotemark{a}} &
\colhead{Seeing} & \colhead{Filter} & \colhead{Inst.} \\
\colhead{(YYYY-mm-dd)} & \colhead{(d)} &
\colhead{(arcsec)} & \colhead{} & \colhead{}
}
\startdata
2012-03-26 & 56012.997 & \sim 0.\arcsec5 & H$\alpha$ & NOT \\
2012-03-27 & 56013.944 & 1.\arcsec5 & grism & SMARTS \\
2012-03-28 & 56014.984 & \nodata & F814M & HST \\
2012-03-30 & 56016.978 & 1.\arcsec5\pm0.25 & B\&C & Bok \\
\enddata
\tablenotetext{a}{At exposure start.}
\tablecomments{The ``C'' command column identifier in the 3 column turns on
math mode for that specific column. One could do the same for the next
column so that dollar signs would not be needed for H$\alpha$
but then all the other text would also be in math mode and thus typeset
in Latin Modern math and you will need to put it back to Roman by hand.
Note that if you do change this column to math mode the dollar signs already
present will not cause a problem. Table \ref{tab:mathmode} is published 
in its entirety in the machine readable format.  A portion is
shown here for guidance regarding its form and content.}
\end{deluxetable*}

\subsection{Tables} \label{subsec:tables}

Tables can be constructed with \latex's standard table environment or the
\aastex's deluxetable environment. The deluxetable construct handles long
tables better but has a larger overhead due to the greater amount of
defined mark up used set up and manipulate the table structure.  The choice
of which to use is up to the author.  Examples of both environments are
used in this manuscript. Table \ref{tab:table} is a simple deluxetable
example that gives the approximate changes in the subscription costs and
author publication charges from 1991 to 2013.

Tables longer than 200 data lines and complex tables should only have a
short example table with the full data set available in the machine
readable format.  The machine readable table will be available in the HTML
version of the article with just a short example in the PDF. Authors are
required to indicate to the reader where the data can be obtained in the
table comments.  Suggested text is given in the comments of Table
\ref{tab:mathmode}.  Authors are encouraged to create their own machine
readable tables using the online tool at
\url{http://authortools.aas.org/MRT/upload.html}.

\aastex\ v6 introduces five new table features that are designed to make
table construction easier and the resulting display better for AAS Journal
authors.  The items are:

\begin{enumerate}
\item Declaring math mode in specific columns,
\item Column decimal alignment, 
\item Automatic column header numbering,
\item Hiding columns, and
\item Splitting wide tables into two or three parts.
\end{enumerate}

Each of these new features are illustrated in following Table examples.
All five features work with the regular \latex\ tabular environment and in
\aastex's deluxetable environment.  The examples in this manuscript also
show where the two process differ.

\subsubsection{Column math mode}

Both the \latex\ tabular and \aastex\ deluxetable require an argument to
define the alignment and number of columns.  The most common values are
``c'', ``l'' and ``r'' for \underline{c}enter, \underline{l}eft, and
\underline{r}ight justification.  If these values are capitalized, e.g.
``C'', ``L'', or ``R'', then that specific column will automatically be in math
mode meaning that \$s are not required.  Note that having embedded dollar
signs in the table does not affect the output.  The third and forth columns
of Table \ref{tab:mathmode} shows how this math mode works.

\subsubsection{Decimal alignment}

Aligning a column by the decimal point can be difficult with only center,
left, and right justification options.  It is possible to use phantom calls
in the data, e.g. {\tt\string\phn}, to align columns by hand but this can
be tedious in long or complex tables.  To address this \aastex\ introduces
the {\tt\string\decimals} command and a new column justification option,
``D'', to align data in that column on the decimal.  In deluxetable the
{\tt\string\decimals} command is invoked before the {\tt\string\startdata}
call but can be anywhere in \latex's tabular environment.

Two other important thing to note when using decimal alignment is that each
decimal column \textit{must end with a space before the ampersand}, e.g.
``\&\&'' is not allowed.  Empty decimal columns are indicated with a decimal,
e.g. ``.''.  Do not use deluxetable's {\tt\string\nodata} command.

The ``D'' alignment token works by splitting the column into two parts on the
decimal.  While this is invisible to the user one must be aware of how it
works so that the headers are accounted for correctly.  All decimal column
headers need to span two columns to get the alignment correct. This can be
done with a multicolumn call, e.g {\tt\string\multicolumn2c\{\}} or
{\tt\string\multicolumn\{2\}\{c\}\{\}}, or use the new
{\tt\string\twocolhead\{\}} command in deluxetable.  Since \latex\ is
splitting these columns into two it is important to get the table width
right so that they appear joined on the page.  You may have to run the
\latex\ compiler twice to get it right.  Table \ref{tab:decimal}
illustrates how decimal alignment works in the tabular environment with a
$\pm$ symbol embedded between the last two columns.

%% Note that the \setcounter and \renewcommand are needed here because
%% this example is using a mix of deluxetable and tabular.  Here the
%% deluxetable counters are set with \tablenum but the situation is a bit
%% more complex for tabular.  Use the first command to set the Table number
%% to ONE LESS than it should be.  The next command will auto increment it
%% to the desired number.
\setcounter{table}{2}
\begin{table}[h!]
\renewcommand{\thetable}{\arabic{table}}
\centering
\caption{Decimal alignment made easy} \label{tab:decimal}
\begin{tabular}{cD@{$\pm$}D}
\tablewidth{0pt}
\hline
\hline
Column & \multicolumn2c{Value} & \multicolumn2c{Uncertainty}\\
\hline
\decimals
A & 1234     & 100.0     \\
B &  123.4   &  10.1     \\
C &  12.34   &   1.01    \\
D &   1.234  &   0.101   \\
E &    .1234 &   0.01001 \\
F &   1.0    &    .      \\
\hline
\multicolumn{5}{c}{NOTE. - Two decimal aligned columns}
\end{tabular}
\end{table}

\subsubsection{Automatic column header numbering} \label{subsubsec:autonumber}

The command {\tt\string\colnumbers} can be included to automatically number
each column as the last row in the header. Per the AAS Journal table format
standards, each column index numbers will be surrounded by parentheses. In
a \latex\ tabular environment the {\tt\string\colnumbers} should be invoked
at the location where the author wants the numbers to appear, e.g. after
the last line of specified table header rows. In deluxetable this command
has to come before {\tt\string\startdata}.  {\tt\string\colnumbers} will
not increment for columns hidden by the ``h'' command, see Section
\ref{subsubsec:hide}.  Table \ref{tab:table} uses this command to
automatically generate column index numbers.

Note that when using decimal alignment in a table the command 
{\tt\string\decimalcolnumbers} must be used instead of 
{\tt\string\colnumbers} and {\tt\string\decimals}. Table \ref{tab:messier}
illustrates this specific functionality.

\subsubsection{Hiding columns} \label{subsubsec:hide}

Entire columns can be \underline{h}idden from display simply by changing
the specified column identifier to ``h''.  In the \latex\ tabular environment
this column identifier conceals the entire column including the header
columns.   In \aastex's deluxetables the header row is specifically
declared with the {\tt\string\tablehead} call and each header column is
marked with {\tt\string\colhead} call.  In order to make a specific header
disappear with the ``h'' column identifier in deluxetable use 
{\tt\string\nocolhead} instead to suppress that particular column header.

Authors can use this option in many different ways.  Since column data can
be easily suppressed authors can include extra information and hid it
based on the comments of co-authors or referees.  For wide tables that will
have a machine readable version, authors could put all the information in
the \latex\ table but use this option to hid as many columns as needed until
it fits on a page. This concealed column table would serve as the
example table for the full machine readable version.  Regardless of how
columns are obscured, authors are responsible for removing any unneeded
column data or alerting the editorial office about how to treat these
columns during production for the final typeset article.

Table \ref{tab:messier} provides some basic information about the first ten
Messier Objects and illustrates how many of these new features can be used
together.  It has automatic column numbering, decimal alignment of the
distances, and one concealed column.  The Common name column
is the third in the \latex\ deluxetable but does not appear when the article
is compiled. This hidden column can be shown simply by changing the ``h'' in
the column identifier preamble to another valid value.  This table also
uses {\tt\string\tablenum} to renumber the table because a \latex\ tabular
table was inserted before it.

\begin{deluxetable*}{cchlDlc}
\tablenum{4}
\tablecaption{Fun facts about the first 10 messier objects\label{tab:messier}}
\tablewidth{0pt}
\tablehead{
\colhead{Messier} & \colhead{NGC/IC} & \nocolhead{Common} & \colhead{Object} &
\multicolumn2c{Distance} & \colhead{} & \colhead{V} \\
\colhead{Number} & \colhead{Number} & \nocolhead{Name} & \colhead{Type} &
\multicolumn2c{(kpc)} & \colhead{Constellation} & \colhead{(mag)}
}
\decimalcolnumbers
\startdata
M1 & NGC 1952 & Crab Nebula & Supernova remnant & 2 & Taurus & 8.4 \\
M2 & NGC 7089 & Messier 2 & Cluster, globular & 11.5 & Aquarius & 6.3 \\
M3 & NGC 5272 & Messier 3 & Cluster, globular & 10.4 & Canes Venatici &  6.2 \\
M4 & NGC 6121 & Messier 4 & Cluster, globular & 2.2 & Scorpius & 5.9 \\
M5 & NGC 5904 & Messier 5 & Cluster, globular & 24.5 & Serpens & 5.9 \\
M6 & NGC 6405 & Butterfly Cluster & Cluster, open & 0.31 & Scorpius & 4.2 \\
M7 & NGC 6475 & Ptolemy Cluster & Cluster, open & 0.3 & Scorpius & 3.3 \\
M8 & NGC 6523 & Lagoon Nebula & Nebula with cluster & 1.25 & Sagittarius & 6.0 \\
M9 & NGC 6333 & Messier 9 & Cluster, globular & 7.91 & Ophiuchus & 8.4 \\
M10 & NGC 6254 & Messier 10 & Cluster, globular & 4.42 & Ophiuchus & 6.4 \\
\enddata
\tablecomments{This table ``hides'' the third column in the \latex\ when compiled.
The Distance is also centered on the decimals.  Note that when using decimal
alignment you need to include the {\tt\string\decimals} command before
{\tt\string\startdata} and all of the values in that column have to have a
space before the next ampersand.}
\end{deluxetable*}

\subsubsection{Splitting a table into multiple horizontal components}

Since the AAS Journals are now all electronic with no print version there is
no reason why tables can not be as wide as authors need them to be.
However, there are some artificial limitations based on the width of a
print page.  The old way around this limitation was to rotate into 
landscape mode and use the smallest available table font
sizes, e.g. {\tt\string\tablewidth}, to get the table to fit.
Unfortunately, this was not alway enough but now along with the hide column
option outlined in Section \ref{subsubsec:hide} there is a new way to break
a table into two or three components so that it flows down a page by
invoking a new table type, splittabular or splitdeluxetable. Within these
tables a new ``B'' column separator is introduced.  Much like the vertical
bar option, ``$\vert$'', that produces a vertical table lines, e.g. Table
\ref{tab:table}, the new ``B'' separator indicates where to \underline{B}reak
a table.  Up to two ``B''s may be included.

Table 5 % \ref{tab:deluxesplit} this freaks it out when it is used!
shows how to split a wide deluxetable into three parts with
the {\tt\string\splitdeluxetable} command.  The {\tt\string\colnumbers}
option is on to show how the automatic column numbering carries through the
second table component, see Section \ref{subsubsec:autonumber}.

The last example, Table \ref{tab:tablesplit}, shows how to split the same
table but with a regular \latex\ tabular call and into two parts. Decimal
alignment is included in the third column and the ``Component'' column is
hidden to illustrate the new features working together.

\begin{splitdeluxetable*}{lccccBcccccBcccc}
\tabletypesize{\scriptsize}
\tablewidth{0pt} 
\tablenum{5}
\tablecaption{Measurements of Emission Lines: two breaks \label{tab:deluxesplit}}
\tablehead{
\colhead{Model} & \colhead{Component}& \colhead{Shift} & \colhead{FWHM} &
\multicolumn{10}{c}{Flux} \\
\colhead{} & \colhead{} & \colhead{($\rm
km~s^{-1}$)}& \colhead{($\rm km~s^{-1}$)} & \multicolumn{10}{c}{($\rm
10^{-17}~erg~s^{-1}~cm^{-2}$)} \\
\cline{5-14}
\colhead{} & \colhead{} &
\colhead{} & \colhead{} & \colhead{Ly$\alpha$} & \colhead{N\,{\footnotesize
V}} & \colhead{Si\,{\footnotesize IV}} & \colhead{C\,{\footnotesize IV}} &
\colhead{Mg\,{\footnotesize II}} & \colhead{H$\gamma$} & \colhead{H$\beta$}
& \colhead{H$\alpha$} & \colhead{He\,{\footnotesize I}} &
\colhead{Pa$\gamma$}
} 
\colnumbers
\startdata 
{       }& BELs& -97.13 &    9117$\pm      38$&    1033$\pm      33$&$< 35$&$<     166$&     637$\pm      31$&    1951$\pm      26$&     991$\pm 30$&    3502$\pm      42$&   20285$\pm      80$&    2025$\pm     116$& 1289$\pm     107$\\ 
{Model 1}& IELs& -4049.123 & 1974$\pm      22$&    2495$\pm      30$&$<     42$&$<     109$&     995$\pm 186$&      83$\pm      30$&      75$\pm      23$&     130$\pm      25$& 357$\pm      94$&     194$\pm      64$& 36$\pm      23$\\
{       }& NELs& \nodata &     641$\pm       4$&     449$\pm 23$&$<      6$&$<       9$&       --            &     275$\pm      18$& 150$\pm      11$&     313$\pm      12$&     958$\pm      43$&     318$\pm 34$& 151$\pm       17$\\
\hline
{       }& BELs& -85 &    8991$\pm      41$& 988$\pm      29$&$<     24$&$<     173$&     623$\pm      28$&    1945$\pm 29$&     989$\pm      27$&    3498$\pm      37$&   20288$\pm      73$& 2047$\pm     143$& 1376$\pm     167$\\
{Model 2}& IELs& -51000 &    2025$\pm      26$& 2494$\pm      32$&$<     37$&$<     124$&    1005$\pm     190$&      72$\pm 28$&      72$\pm      21$&     113$\pm      18$&     271$\pm      85$& 205$\pm      72$& 34$\pm      21$\\
{       }& NELs& 52 &     637$\pm      10$&     477$\pm 17$&$<      4$&$<       8$&       --            &     278$\pm      17$& 153$\pm      10$&     317$\pm      15$&     969$\pm      40$&     325$\pm 37$&
     147$\pm       22$\\
\enddata
\tablecomments{This is an example of how to split a deluxetable. You can
split any table with this command into two or three parts.  The location of
the split is given by the author based on the placement of the ``B''
indicators in the column identifier preamble.  For more information please
look at the new \aastex\ instructions.}
\end{splitdeluxetable*}

%\clearpage

\setcounter{table}{5}
\begin{table*}[h!]
\renewcommand{\thetable}{\arabic{table}}
\caption{Measurements of Emission Lines: one break\label{tab:tablesplit}}
\begin{splittabular}{lhDccccBccccccc}
%\multicolumn{5}{c}{Table 6} \\
%\multicolumn{5}{c}{Measurements of Emission Lines} \\
\hline 
\hline 
Model & Component & \multicolumn2c{Shift} & FWHM & 
\multicolumn{10}{c}{Flux} \\
 & & \multicolumn2c{($\rm km~s^{-1}$)} & {($\rm km~s^{-1}$)} & 
\multicolumn{10}{c}{($\rm 10^{-17}~erg~s^{-1}~cm^{-2}$)} \\
\cline{5-15}
 & & & & & {Ly$\alpha$} & {N\,{\footnotesize V}} & 
{Si\,{\footnotesize IV}} & {C\,{\footnotesize IV}} &
{Mg\,{\footnotesize II}} & {H$\gamma$} & {H$\beta$}
& {H$\alpha$} & {He\,{\footnotesize I}} & {Pa$\gamma$} \\
%\hline
\decimalcolnumbers
 & BELs& -97.13 &    9117$\pm      38$&    1033$\pm      33$&$< 35$&$<     166$&     637$\pm      31$&    1951$\pm      26$&     991$\pm 30$&    3502$\pm      42$&   20285$\pm      80$&    2025$\pm     116$& 1289$\pm     107$\\
Model 1 & IELs& -4049.123 & 1974$\pm      22$&    2495$\pm      30$&$<     42$&$<     109$&     995$\pm 186$&      83$\pm      30$&      75$\pm      23$&     130$\pm      25$& 357$\pm      94$&     194$\pm      64$& 36$\pm      23$\\
 & NELs& . &     641$\pm       4$&     449$\pm 23$&$<      6$&$<       9$&       --            &     275$\pm      18$& 150$\pm      11$&     313$\pm      12$&     958$\pm      43$&     318$\pm 34$& 151$\pm       17$\\
\hline
 & BELs& -85 &    8991$\pm      41$& 988$\pm      29$&$<     24$&$<     173$&     623$\pm      28$&    1945$\pm 29$&     989$\pm      27$&    3498$\pm      37$&   20288$\pm      73$& 2047$\pm     143$& 1376$\pm     167$\\
Model 2 & IELs& -51000 &    2025$\pm      26$& 2494$\pm      32$&$<     37$&$<     124$&    1005$\pm     190$&      72$\pm 28$&      72$\pm      21$&     113$\pm      18$&     271$\pm      85$& 205$\pm      72$& 34$\pm      21$\\
 & NELs& 52 &     637$\pm      10$&     477$\pm 17$&$<      4$&$<       8$&       --            &     278$\pm      17$& 153$\pm      10$&     317$\pm      15$&     969$\pm      40$&     325$\pm 37$& 147$\pm       22$\\
\hline
\end{splittabular}
\end{table*}

\subsection{Figures\label{subsec:figures}}

%% The "ht!" tells LaTeX to put the figure "here" first, at the "top" next
%% and to override the normal way of calculating a float position
%\begin{figure}[ht!]
%\plotone{cost-eps-converted-to.pdf}
%\caption{The subscription and author publication costs from 1991 to 2013.
%The data comes from Table \ref{tab:table}.\label{fig:general}}
%\end{figure}

Authors can include a wide number of different graphics with their articles
in  encapsulated postscript (EPS) or portable document format (PDF).  These
range from general figures all authors are familiar with to new enhanced
graphics that can only be fully experienced in HTML.  The later include
animations, figure sets and interactive figures.  This portion of the
article provides examples for setting up all these graphics in with the
latest version of \aastex.

\subsection{General figures\label{subsec:general}}

\aastex\ has a {\tt\string\plotone} command to display a figure
consisting of one EPS/PDF file.  Figure \ref{fig:general} is an example
which uses the data from Table \ref{tab:table}. For a general figure
consisting of two EPS/PDF files the {\tt\string\plottwo} command can be
used to position the two image files side by side.  Figure \ref{fig:f2}
shows the Swift/XRT X-ray light curves of two recurrent novae.  The data
from Figures \ref{fig:f2} through \ref{fig:fig4} are taken from Table 2 of
\citet{2011ApJS..197...31S}.

%\begin{figure}
%\plottwo{RS_Oph-eps-converted-to.pdf}{U_Sco-eps-converted-to.pdf}
%\caption{Swift/XRT X-ray light curves of RS Oph and U Sco which represent
%the two canonical recurrent types, a long period system with a red giant
%secondary and a short period system with a dwarf/sub-dwarf secondary,
%respectively.\label{fig:f2}}
%\end{figure}

Both {\tt\string\plotone} and {\tt\string\plottwo} take a
{\tt\string\caption} and an optional {\tt\string\figurenum} command to
specify the figure number\footnote{It is better to not use
{\tt\string\figurenum} and let LaTeX auto-increment all the figures. If you
do use this command you need to mark all of them accordingly.}.  Each is
based on the {\tt\string graphicx} package command,
{\tt\string\includegraphics}.  Authors are welcome to use
{\tt\string\includegraphics} along with its optional arguments that control
the height, width, scale, and position angle of a file within the figure.
More information on the full usage of {\tt\string\includegraphics} can be
found at \break
\url{https://en.wikibooks.org/wiki/LaTeX/Importing\_Graphics\#Including\_graphics}.

\subsection{Grid figures}

Including more than two EPS/PDF files in a single figure call can be tricky
easily format.  To make the process easier for authors \aastex\ v6 offers
{\tt\string\gridline} which allows any number of individual EPS/PDF file
calls within a single figure.  Each file cited in a {\tt\string\gridline}
will be displayed in a row.  By adding more {\tt\string\gridline} calls an
author can easily construct a matrix X by Y individual files as a
single general figure.

For each {\tt\string\gridline} command a EPS/PDF file is called by one of
four different commands.  These are {\tt\string\fig},
{\tt\string\rightfig}, {\tt\string\leftfig}, and {\tt\string\boxedfig}.
The first file call specifies no image position justification while the
next two will right and left justify the image, respectively.  The
{\tt\string\boxedfig} is similar to {\tt\string\fig} except that a box is
drawn around the figure file when displayed. Each of these commands takes
three arguments.  The first is the file name.  The second is the width that
file should be displayed at.  While any natural \latex\ unit is allowed, it
is recommended that author use fractional units with the
{\tt\string\textwidth}.  The last argument is text for a subcaption.

Figure \ref{fig:pyramid} shows an inverted pyramid of individual
figure constructed with six individual EPS files using the
{\tt\string\gridline} option.

%\begin{figure*}
%\gridline{\fig{V2491_Cyg-eps-converted-to.pdf}{0.3\textwidth}{(a)}
%          \fig{CSS081007-eps-converted-to.pdf}{0.3\textwidth}{(b)}
%          \fig{LMC_2009-eps-converted-to.pdf}{0.3\textwidth}{(c)}
%          }
%\gridline{\fig{RS_Oph-eps-converted-to.pdf}{0.3\textwidth}{(d)}
%          \fig{U_Sco-eps-converted-to.pdf}{0.3\textwidth}{(e)}
%          }
%\gridline{\fig{KT_Eri-eps-converted-to.pdf}{0.3\textwidth}{(f)}}
%\caption{Inverted pyramid figure of six individual files. The nova are
%(a) V2491 Cyg, (b) HV Cet, (c) LMC 2009, (d) RS Oph, (e) U Sco, and
%(f) KT Eri.\label{fig:pyramid}}
%\end{figure*}

\subsection{Figure sets}

A large collection of similar style figures should be grouped together as a
figure set.  The derived PDF article will only shows an example figure
while the enhanced content is available in the figure set in the electronic
edition.  The advantage of a figure set gives the reader the ability to
easily sort through the figure collection to find individual component
figures.  All of the figure set components, along with their html framework,
are also available for download in a .tar.gz package.

Special \latex\ mark up is required to create a figure set.  Prior to
\aastex\ v6 the underlying mark up commands had to be inserted by hand
but is now included.  Note that when an article with figure set is compiled
in \latex\ none of the component figures are shown and a floating Figure
Set caption will appear in the resulting PDF.

\figsetstart
\figsetnum{4}
\figsettitle{Swift X-ray light curves}

\figsetgrpstart
\figsetgrpnum{1.1}
\figsetgrptitle{KT Eri}
\figsetplot{KT_Eri-eps-converted-to.pdf}
\figsetgrpnote{The Swift/XRT X-ray light curve for the first year after
outburst.}
\figsetgrpend

\figsetgrpstart
\figsetgrpnum{1.2}
\figsetgrptitle{RS Oph}
\figsetplot{RS_Oph-eps-converted-to.pdf}
\figsetgrpnote{The Swift/XRT X-ray light curve for the first year after
outburst.}
\figsetgrpend

\figsetgrpstart
\figsetgrpnum{1.3}
\figsetgrptitle{U Sco}
\figsetplot{U_Sco-eps-converted-to.pdf}
\figsetgrpnote{The Swift/XRT X-ray light curve for the first year after
outburst.}
\figsetgrpend

\figsetgrpstart
\figsetgrpnum{1.4}
\figsetgrptitle{V2491 Cyg}
\figsetplot{V2491_Cyg-eps-converted-to.pdf}
\figsetgrpnote{The Swift/XRT X-ray light curve for the first year after
outburst.}
\figsetgrpend

\figsetgrpstart
\figsetgrpnum{1.5}
\figsetgrptitle{Nova LMC 2009}
\figsetplot{LMC_2009-eps-converted-to.pdf}
\figsetgrpnote{The Swift/XRT X-ray light curve for the first year after
outburst.}
\figsetgrpend

\figsetgrpstart
\figsetgrpnum{1.6}
\figsetgrptitle{HV Cet}
\figsetplot{CSS081007-eps-converted-to.pdf}
\figsetgrpnote{The Swift/XRT X-ray light curve for the first year after
outburst.}
\figsetgrpend

\figsetgrpstart
\figsetgrpnum{1.7}
\figsetgrptitle{V2672 Oph}
\figsetplot{V2672_Oph-eps-converted-to.pdf}
\figsetgrpnote{The Swift/XRT X-ray light curve for the first year after
outburst.}
\figsetgrpend

\figsetgrpstart
\figsetgrpnum{1.8}
\figsetgrptitle{V407 Cyg}
\figsetplot{V407_Cyg-eps-converted-to.pdf}
\figsetgrpnote{The Swift/XRT X-ray light curve for the first year after
outburst.}
\figsetgrpend

\figsetend

%\begin{figure}
%\plotone{KT_Eri-eps-converted-to.pdf}
%\caption{The Swift/XRT X-ray light curve for the first year after
%outburst of the suspected recurrent nova KT Eri. At a maximum count rate of 
%328 ct/s, KT Eri was the brightest nova in X-rays observed to date. All 
%the component figures are available in the Figure Set. \label{fig:fig4}}
%\end{figure}

Authors are encouraged to use an online tool at
\url{http://authortools.aas.org/FIGSETS/make-figset.html} to generate their
own specific figure set mark up to incorporate into their \latex\ articles.

\subsection{Animations}

Authors may include animations in their articles.  A single still frame from 
the animation should be included as a regular figure to serve as an example.
The associated figure caption should indicate to the reader exactly what the
animation shows and that the animation is available online.

%\begin{figure}
%\plotone{video3-eps-converted-to.pdf}
%\caption{Example image from the animation which is available in the electronic
%edition.}
%\end{figure}

\subsection{Interactive figures}

Interactive figures give the reader the ability to manipulate the
information contained in an image which can add clarity or help further the
author's narrative.  These figures consist of two parts, the figure file in
a specific format and a javascript and html frame work that provides the
interactive control.  An example of an interactive figure is a 3D model.
The underlying figure is a X3D file while x3dom.js is the javascript driver
that displays it. An author created interface is added via a html wrapper.
The first 3D model published by the AAS Journals using this technique was
\citet{2014ApJ...793..127V}.  Authors should consult the online tutorials
for more information on how to construct their own interactive figures.

As with animations authors should include a non-interactive regular figure
to use as an example.  The example figure should also indicate to the reader
that the enhanced figure is interactive and can be accessed online.

\section{Displaying mathematics} \label{sec:displaymath}

The most common mathematical symbols and formulas are in the amsmath
package.  \aastex\ requires this package so there is no need to
specifically call for it in the document preamble.  Most modern \latex\
distributions already contain this package.  If you do not have this
package or the other required packages, revtex4-1, latexsym, graphicx,
amssymb, longtable, and epsf, they can be obtained from 
\url{http://www.ctan.org}

Mathematics can be displayed either within the text, e.g. $E = mc^2$, or
separate from in an equation.  In order to be properly rendered, all inline
math text has to be declared by surrounding the math by dollar signs (\$).

A complex equation example with inline math as part of the explanation
follows.

\begin{equation}
\bar v(p_2,\sigma_2)P_{-\tau}\hat a_1\hat a_2\cdots
\hat a_nu(p_1,\sigma_1) ,
\end{equation}
where $p$ and $\sigma$ label the initial $e^{\pm}$ four-momenta
and helicities $(\sigma = \pm 1)$, $\hat a_i=a^\mu_i\gamma_\nu$
and $P_\tau=\frac{1}{2}(1+\tau\gamma_5)$ is a chirality projection
operator $(\tau = \pm1)$.  This produces a single line formula.  \latex\ will
auto-number this and any subsequent equations.  If no number is desired then
the {\tt\string equation} call should be replaced with {\tt\string displaymath}.

\latex\ can also handle a a multi-line equation.  Use {\tt\string eqnarray}
for more than one line and end each line with a
\textbackslash\textbackslash.  Each line will be numbered unless the
\textbackslash\textbackslash\ is preceded by a {\tt\string\nonumber}
command.  Alignment points can be added with ampersands (\&).  There should be
two ampersands per line. In the examples they are centered on the equal
symbol.

\begin{eqnarray}
\gamma^\mu  & = &
 \left(
\begin{array}{cc}
0 & \sigma^\mu_+ \\
\sigma^\mu_- & 0
\end{array}     \right) ,
 \gamma^5= \left(
\begin{array}{cc}
-1 &   0\\
0 &   1
\end{array}     \right)  , \\
\sigma^\mu_{\pm}  & = &   ({\bf 1} ,\pm \sigma) , 
\end{eqnarray}

\begin{eqnarray}
\hat a & = & \left(
\begin{array}{cc}
0 & (\hat a)_+\\
(\hat a)_- & 0
\end{array}\right), \nonumber \\
(\hat a)_\pm & = & a_\mu\sigma^\mu_\pm 
\end{eqnarray}

%% Putting eqnarrays or equations inside the mathletters environment groups
%% the enclosed equations by letter. For instance, the eqnarray below, instead
%% of being numbered, say, (4) and (5), would be numbered (4a) and (4b).
%% LaTeX the paper and look at the output to see the results.

\section{Revision tracking and color highlighting} \label{sec:highlight}

Authors sometimes use color to highlight changes to their manuscript in
response to editor and referee comments.  In \aastex\ new commands
have been introduced to make this easier and formalize the process.

The first method is through a new set of editing mark up commands that
specifically identify what has been changed.  These commands are
{\tt\string\added\{<text>\}}, {\tt\string\deleted\{<text>\}}, and
{\tt\string\replaced\{<old text>\}\{<replaced text>\}}. To activate these
commands the {\tt\string trackchanges} option must be used in the
{\tt\string\documentclass} call.  When compiled this will produce the
marked text in red.  The {\tt\string\explain\{<text>\}} can be used to add
text to provide information to the reader describing the change.  Its
output is purple italic font. To see how {\tt\string\added\{<important
added info>\}}, {\tt\string\deleted\{<this can be deleted text>\}},
{\tt\string\replaced\{<old data>\}\{<replaced data>\}}, and \break
{\tt\string\explain\{<text explaining the change>\}} commands will produce
\added{important added information}\deleted{, deleted text, and }
\replaced{old data}{and replaced data,} toggle between versions compiled with
and without the {\tt\string trackchanges} option.\explain{text explaining
the change}

A summary list of all these tracking commands can be produced at the end of
the article by adding the {\tt\string\listofchanges} just before the
{\tt\string\end\{document\}} call.  The page number for each change will be
provided. If the {\tt\string linenumbers} option is also included in the
documentcall call then not only will all the lines in the article be
numbered for handy reference but the summary list will also include the
line number for each change.

The second method does not have the ability to highlight the specific
nature of the changes but does allow the author to document changes over
multiple revisions.  The commands are {\tt\string\edit1\{<text>\}},
{\tt\string\edit2\{<text>\}} and {\tt\string\edit3\{<text>\}} and they
produce {\tt\string<text>} that is highlighted in bold red, italic blue and
underlined purple, respectively.  Authors should use the first command to
\edit1{indicated which text has been changed from the first revision.}  The
second command is to highlight \edit2{new or modified text from a second
revision}.  If a third revision is needed then the last command should be used 
\edit3{to show this changed text}.  Since over 90\% of all manuscripts are
accepted after the 3rd revision these commands make it easy to identify
what text has been added and when.  Once the article is accepted all the
highlight color can be turned off simply by adding the
{\tt\string\turnoffediting} command in the preamble. Likewise, the new commands
{\tt\string\turnoffeditone}, {\tt\string\turnoffedittwo}, and
{\tt\string\turnoffeditthree} can be used to only turn off the 
{\tt\string\edit1\{<text>\}}, {\tt\string\edit2\{<text>\}} and 
{\tt\string\edit3\{<text>\}}, respectively.

Similar to marking editing changes with the {\tt\string\edit} options there
are also the {\tt\string\authorcomments1\{<text>\}}, 
{\tt\string\authorcomments2\{<text>\}} and
{\tt\string\authorcomments3\{<text>\}} commands.  These produce the same
bold red, italic blue and underlined purple text but when the
{\tt\string\turnoffediting} command is present the {\tt\string<text>}
material does not appear in the manuscript.  Authors can use these commands
to mark up text that they are not sure should appear in the final
manuscript or as a way to communicate comments between co-authors when
writing the article.

\section{Software and third party data repository citations} \label{sec:cite}

The AAS Journals would like to encourage authors to change software and
third party data repository references from the current standard of a
footnote to a first class citation in the bibliography.  As a bibliographic
citation these important references will be more easily captured and credit
will be given to the appropriate people.

The first step to making this happen is to have the data or software in
a long term repository that has made these items available via a persistent
identifier like a Digital Object Identifier (DOI).  A list of repositories
that satisfy this criteria plus each one's pros and cons are given at \break
\url{https://github.com/AASJournals/Tutorials/tree/master/Repositories}.

In the bibliography the format for data or code follows this format: \\

\noindent author year, title, version, publisher, prefix:identifier\\

\citet{2015ApJ...805...23C} provides a example of how the citation in the
article references the external code at
\doi{10.5281/zenodo.15991}.  Unfortunately, bibtex does
not have specific bibtex entries for these types of references so the
``@misc'' type should be used.  The Repository tutorial explains how to
code the ``@misc'' type correctly.  The most recent aasjournal.bst file,
available with \aastex\ v6, will output bibtex ``@misc'' type properly.

%% If you wish to include an acknowledgments section in your paper,
%% separate it off from the body of the text using the \acknowledgments
%% command.
\acknowledgments

We thank all the people that have made this AASTeX what it is today.  This
includes but not limited to Bob Hanisch, Chris Biemesderfer, Lee Brotzman,
Pierre Landau, Arthur Ogawa, Maxim Markevitch, Alexey Vikhlinin and Amy
Hendrickson. Also special thanks to David Hogg and Daniel Foreman-Mackey
for the new "modern" style design. Considerable help was provided via bug
reports and hacks from numerous people including Patricio Cubillos, Alex
Drlica-Wagner, Sean Lake, Michele Bannister, Peter Williams, and Jonathan
Gagne.

%% To help institutions obtain information on the effectiveness of their 
%% telescopes the AAS Journals has created a group of keywords for telescope 
%% facilities.
%
%% Following the acknowledgments section, use the following syntax and the
%% \facility{} or \facilities{} macros to list the keywords of facilities used 
%% in the research for the paper.  Each keyword is check against the master 
%% list during copy editing.  Individual instruments can be provided in 
%% parentheses, after the keyword, but they are not verified.

\vspace{5mm}
\facilities{HST(STIS), Swift(XRT and UVOT), AAVSO, CTIO:1.3m,
CTIO:1.5m,CXO}

%% Similar to \facility{}, there is the optional \software command to allow 
%% authors a place to specify which programs were used during the creation of 
%% the manusscript. Authors should list each code and include either a
%% citation or url to the code inside ()s when available.

\software{astropy \citep{2013A&A...558A..33A},  
          Cloudy \citep{2013RMxAA..49..137F}, 
          SExtractor \citep{1996A&AS..117..393B}
          }

%% Appendix material should be preceded with a single \appendix command.
%% There should be a \section command for each appendix. Mark appendix
%% subsections with the same markup you use in the main body of the paper.

%% Each Appendix (indicated with \section) will be lettered A, B, C, etc.
%% The equation counter will reset when it encounters the \appendix
%% command and will number appendix equations (A1), (A2), etc. The
%% Figure and Table counter will not reset.

\appendix

\section{Appendix information}

Appendices can be broken into separate sections just like in the main text.
The only difference is that each appendix section is indexed by a letter
(A, B, C, etc.) instead of a number.  Likewise numbered equations have
the section letter appended.  Here is an equation as an example.

\begin{equation}
I = \frac{1}{1 + d_{1}^{P (1 + d_{2} )}}
\end{equation}

Appendix tables and figures should not be numbered like equations. Instead
they should continue the sequence from the main article body.

\section{Author publication charges} \label{sec:pubcharge}

Finally some information about the AAS Journal's publication charges.
In April 2011 the traditional way of calculating author charges based on 
the number of printed pages was changed.  The reason for the change
was due to a recognition of the growing number of article items that could not 
be represented in print. Now author charges are determined by a number of
digital ``quanta''.  A single quantum is 350 words, one figure, one table,
and one enhanced digital item.  For the latter this includes machine readable
tables, figure sets, animations, and interactive figures.  The current cost
is \$27 per word quantum and \$30 for all other quantum type.

\section{Rotating tables} \label{sec:rotate}

The process of rotating tables into landscape mode is slightly different in
\aastex v6.2. Instead of the {\tt\string\rotate} command, a new environment
has been created to handle this task. To place a single page table in a
landscape mode start the table portion with
{\tt\string\begin\{rotatetable\}} and end with
{\tt\string\end\{rotatetable\}}.

Tables that exceed a print page take a slightly different environment since
both rotation and long table printing are required. In these cases start
with {\tt\string\begin\{longrotatetable\}} and end with
{\tt\string\end\{longrotatetable\}}. Table \ref{chartable} is an
example of a multi-page, rotated table.

\begin{longrotatetable}
\begin{deluxetable*}{lllrrrrrrll}
\tablecaption{Observable Characteristics of 
Galactic/Magellanic Cloud novae with X-ray observations\label{chartable}}
\tablewidth{700pt}
\tabletypesize{\scriptsize}
\tablehead{
\colhead{Name} & \colhead{V$_{max}$} & 
\colhead{Date} & \colhead{t$_2$} & 
\colhead{FWHM} & \colhead{E(B-V)} & 
\colhead{N$_H$} & \colhead{Period} & 
\colhead{D} & \colhead{Dust?} & \colhead{RN?} \\ 
\colhead{} & \colhead{(mag)} & \colhead{(JD)} & \colhead{(d)} & 
\colhead{(km s$^{-1}$)} & \colhead{(mag)} & \colhead{(cm$^{-2}$)} &
\colhead{(d)} & \colhead{(kpc)} & \colhead{} & \colhead{}
} 
\startdata
CI Aql & 8.83 (1) & 2451665.5 (1) & 32 (2) & 2300 (3) & 0.8$\pm0.2$ (4) & 1.2e+22 & 0.62 (4) & 6.25$\pm5$ (4) & N & Y \\
{\bf CSS081007} & \nodata & 2454596.5 & \nodata & \nodata & 0.146 & 1.1e+21 & 1.77 (5) & 4.45$\pm1.95$ (6) & \nodata & \nodata \\
GQ Mus & 7.2 (7) & 2445352.5 (7) & 18 (7) & 1000 (8) & 0.45 (9) & 3.8e+21  & 0.059375 (10) & 4.8$\pm1$ (9) & N (7) & \nodata \\
IM Nor & 7.84 (11) & 2452289 (2) & 50 (2) & 1150 (12) & 0.8$\pm0.2$ (4) & 8e+21 & 0.102 (13) & 4.25$\pm3.4$ (4) & N & Y \\
{\bf KT Eri} & 5.42 (14) & 2455150.17 (14) & 6.6 (14) & 3000 (15) & 0.08 (15) & 5.5e+20 & \nodata & 6.5 (15) & N & M \\
{\bf LMC 1995} & 10.7 (16) & 2449778.5 (16) & 15$\pm2$ (17) & \nodata & 0.15 (203) & 7.8e+20  & \nodata & 50 & \nodata & \nodata \\
LMC 2000 & 11.45 (18) & 2451737.5 (18) & 9$\pm2$ (19) & 1700 (20) & 0.15 (203) & 7.8e+20  & \nodata & 50 & \nodata & \nodata \\
{\bf LMC 2005} & 11.5 (21) & 2453700.5 (21) & 63 (22) & 900 (23) & 0.15 (203) & 1e+21 & \nodata & 50  & M (24) & \nodata \\
{\bf LMC 2009a} & 10.6 (25) & 2454867.5 (25) & 4$\pm1$  & 3900 (25) & 0.15 (203)  & 5.7e+20 & 1.19 (26) & 50 & N & Y \\
{\bf SMC 2005} & 10.4 (27) & 2453588.5 (27) & \nodata & 3200 (28) & \nodata & 5e+20  & \nodata & 61 & \nodata & \nodata \\
{\bf QY Mus} & 8.1 (29) & 2454739.90 (29) & 60:  & \nodata & 0.71 (30) & 4.2e+21  & \nodata & \nodata & M & \nodata \\
{\bf RS Oph} & 4.5 (31) & 2453779.44 (14) & 7.9 (14) & 3930 (31) & 0.73 (32) & 2.25e+21 & 456 (33) & 1.6$\pm0.3$ (33) & N (34) & Y \\
{\bf U Sco} & 8.05 (35) & 2455224.94 (35) & 1.2 (36) & 7600 (37) & 0.2$\pm0.1$ (4) & 1.2e+21 & 1.23056 (36) & 12$\pm2$ (4) & N & Y \\
{\bf V1047 Cen} & 8.5 (38) & 2453614.5 (39) & 6 (40) & 840 (38) & \nodata & 1.4e+22  & \nodata & \nodata & \nodata & \nodata \\
{\bf V1065 Cen} & 8.2 (41) & 2454123.5 (41) & 11 (42) & 2700 (43) & 0.5$\pm0.1$ (42) & 3.75e+21 & \nodata & 9.05$\pm2.8$ (42) & Y (42) & \nodata \\
V1187 Sco & 7.4 (44) & 2453220.5 (44) & 7: (45) & 3000 (44) & 1.56 (44) & 8.0e+21 & \nodata & 4.9$\pm0.5$ (44) & N & \nodata \\
{\bf V1188 Sco} & 8.7 (46) & 2453577.5 (46) & 7 (40) & 1730 (47) & \nodata & 5.0e+21  & \nodata & 7.5 (39) & \nodata & \nodata \\
{\bf V1213 Cen} & 8.53 (48) & 2454959.5 (48) & 11$\pm2$ (49) & 2300 (50) & 2.07 (30) & 1.0e+22 & \nodata & \nodata & \nodata & \nodata \\
{\bf V1280 Sco} & 3.79 (51) & 2454147.65 (14) & 21 (52) & 640 (53) & 0.36 (54) & 1.6e+21  & \nodata & 1.6$\pm0.4$ (54) & Y (54) & \nodata \\
{\bf V1281 Sco} & 8.8 (55) & 2454152.21 (55) & 15:& 1800 (56) & 0.7 (57) & 3.2e+21 & \nodata & \nodata & N & \nodata \\
{\bf V1309 Sco} & 7.1 (58) & 2454714.5 (58) & 23$\pm2$ (59) & 670 (60) & 1.2 (30) & 4.0e+21 & \nodata & \nodata & \nodata & \nodata \\
{\bf V1494 Aql} & 3.8 (61) & 2451515.5 (61) & 6.6$\pm0.5$ (61) & 1200 (62) & 0.6 (63) & 3.6e+21  & 0.13467 (64) & 1.6$\pm0.1$ (63) & N & \nodata \\
{\bf V1663 Aql} & 10.5 (65) & 2453531.5 (65) & 17 (66) & 1900 (67) & 2: (68) & 1.6e+22  & \nodata & 8.9$\pm3.6$ (69) & N & \nodata \\
V1974 Cyg & 4.3 (70) & 2448654.5 (70) & 17 (71) & 2000 (19) & 0.36$\pm0.04$ (71) & 2.7e+21  & 0.081263 (70) & 1.8$\pm0.1$ (72) & N & \nodata \\
{\bf V2361 Cyg} & 9.3 (73) & 2453412.5 (73) & 6 (40) & 3200 (74) & 1.2: (75) & 7.0e+21 & \nodata & \nodata & Y (40) & \nodata \\
{\bf V2362 Cyg} & 7.8 (76) & 2453831.5 (76) & 9 (77) & 1850 (78) & 0.575$\pm0.015$ (79) & 4.4e+21  & 0.06577 (80) & 7.75$\pm3$ (77) & Y (81) & \nodata \\
{\bf V2467 Cyg} & 6.7 (82) & 2454176.27 (82) & 7 (83) & 950 (82) & 1.5 (84) & 1.4e+22  & 0.159 (85) & 3.1$\pm0.5$ (86) & M (87) & \nodata \\
{\bf V2468 Cyg} & 7.4 (88) & 2454534.2 (88) & 10: & 1000 (88) & 0.77 (89) & 1.0e+22  & 0.242 (90) & \nodata & N & \nodata \\
{\bf V2491 Cyg} & 7.54 (91) & 2454567.86 (91) & 4.6 (92) & 4860 (93) & 0.43 (94) & 4.7e+21  & 0.09580: (95) & 10.5 (96) & N & M \\
V2487 Oph & 9.5 (97) & 2450979.5 (97) & 6.3 (98) & 10000 (98) & 0.38$\pm0.08$ (98) & 2.0e+21 & \nodata & 27.5$\pm3$ (99) & N (100) & Y (101) \\
{\bf V2540 Oph} & 8.5 (102) & 2452295.5 (102) & \nodata & \nodata & \nodata & 2.3e+21 & 0.284781 (103) & 5.2$\pm0.8$ (103) & N & \nodata \\
V2575 Oph & 11.1 (104) & 2453778.8 (104) & 20: & 560 (104) & 1.4 (105) & 3.3e+21 & \nodata & \nodata & N (105) & \nodata \\
{\bf V2576 Oph} & 9.2 (106) & 2453832.5 (106) & 8: & 1470 (106) & 0.25 (107) & 2.6e+21  & \nodata & \nodata & N & \nodata \\
{\bf V2615 Oph} & 8.52 (108) & 2454187.5 (108) & 26.5 (108) & 800 (109) & 0.9 (108) & 3.1e+21  & \nodata & 3.7$\pm0.2$ (108) & Y (110) & \nodata \\
{\bf V2670 Oph} & 9.9 (111) & 2454613.11 (111) & 15: & 600 (112) & 1.3: (113) & 2.9e+21  & \nodata & \nodata & N (114) & \nodata \\
{\bf V2671 Oph} & 11.1 (115) & 2454617.5 (115) & 8: & 1210 (116) & 2.0 (117) & 3.3e+21  & \nodata & \nodata & M (117) & \nodata \\
{\bf V2672 Oph} & 10.0 (118) & 2455060.02 (118) & 2.3 (119) & 8000 (118) & 1.6$\pm0.1$ (119) & 4.0e+21  & \nodata & 19$\pm2$ (119) & \nodata & M \\
V351 Pup & 6.5 (120) & 2448617.5 (120) & 16 (121) & \nodata & 0.72$\pm0.1$ (122) & 6.2e+21 & 0.1182 (123) & 2.7$\pm0.7$ (122) & N & \nodata \\
{\bf V382 Nor} & 8.9 (124) & 2453447.5 (124) & 12 (40) & 1850 (23) & \nodata & 1.7e+22 & \nodata & \nodata & \nodata & \nodata \\
V382 Vel & 2.85 (125) & 2451320.5 (125) & 4.5 (126) & 2400 (126) & 0.05: (126) & 3.4e+21  & 0.146126 (127) & 1.68$\pm0.3$ (126) & N & \nodata \\
{\bf V407 Cyg} & 6.8 (128) & 2455266.314 (128) & 5.9 (129) & 2760 (129) & 0.5$\pm0.05$ (130) & 8.8e+21 & 15595 (131) & 2.7 (131) & \nodata & Y \\
{\bf V458 Vul} & 8.24 (132) & 2454322.39 (132) & 7 (133) & 1750 (134) & 0.6 (135) & 3.6e+21 & 0.06812255 (136) & 8.5$\pm1.8$ (133) & N (135) & \nodata \\
{\bf V459 Vul} & 7.57 (137) & 2454461.5 (137) & 18 (138) & 910 (139) & 1.0 (140) & 5.5e+21  & \nodata & 3.65$\pm1.35$ (138) & Y (140) & \nodata \\
V4633 Sgr & 7.8 (141) & 2450895.5 (141) & 19$\pm3$ (142) & 1700 (143) & 0.21 (142) & 1.4e+21  & 0.125576 (144) & 8.9$\pm2.5$ (142) & N & \nodata \\
{\bf V4643 Sgr} & 8.07 (145) & 2451965.867 (145) & 4.8 (146) & 4700 (147) & 1.67 (148) & 1.4e+22 & \nodata & 3 (148) & N & \nodata \\
{\bf V4743 Sgr} & 5.0 (149) & 2452537.5 (149) & 9 (150) & 2400 (149) & 0.25 (151) & 1.2e+21 & 0.281 (152) & 3.9$\pm0.3$ (151) & N & \nodata \\
{\bf V4745 Sgr} & 7.41 (153) & 2452747.5 (153) & 8.6 (154) & 1600 (155) & 0.1 (154) & 9.0e+20  & 0.20782 (156) & 14$\pm5$ (154) & \nodata & \nodata \\
{\bf V476 Sct} & 10.3 (157) & 2453643.5 (157) & 15 (158) & \nodata & 1.9 (158) & 1.2e+22  & \nodata & 4$\pm1$ (158) & M (159) & \nodata \\
{\bf V477 Sct} & 9.8 (160) & 2453655.5 (160) & 3 (160) & 2900 (161) & 1.2: (162) & 4e+21  & \nodata & \nodata & M (163) & \nodata \\
{\bf V5114 Sgr} & 8.38 (164) & 2453081.5 (164) & 11 (165) & 2000 (23) & \nodata & 1.5e+21  & \nodata & 7.7$\pm0.7$ (165) & N (166) & \nodata \\
{\bf V5115 Sgr} & 7.7 (167) & 2453459.5 (167) & 7 (40) & 1300 (168) & 0.53 (169) & 2.3e+21  & \nodata & \nodata & N (169) & \nodata \\
{\bf V5116 Sgr} & 8.15 (170) & 2453556.91 (170) & 6.5 (171) & 970 (172) & 0.25 (173) & 1.5e+21 & 0.1238 (171) & 11$\pm3$ (173) & N (174) & \nodata \\
{\bf V5558 Sgr} & 6.53 (175) & 2454291.5 (175) & 125 (176) & 1000 (177) & 0.80 (178) & 1.6e+22  & \nodata & 1.3$\pm0.3$ (176) & N (179) & \nodata \\
{\bf V5579 Sgr} & 5.56 (180) & 2454579.62 (180) & 7: & 1500 (23) & 1.2 (181) & 3.3e+21 & \nodata & \nodata & Y (181) & \nodata \\
{\bf V5583 Sgr} & 7.43 (182) & 2455051.07 (182) & 5: & 2300 (182) & 0.39 (30) & 2.0e+21 & \nodata & 10.5 & \nodata & \nodata \\
{\bf V574 Pup} & 6.93 (183) & 2453332.22 (183) & 13 (184) & 2800 (184) & 0.5$\pm0.1$  & 6.2e+21 & \nodata & 6.5$\pm1$  & M (185) & \nodata \\
{\bf V597 Pup} & 7.0 (186) & 2454418.75 (186) & 3: & 1800 (187) & 0.3 (188) & 5.0e+21  & 0.11119 (189) & \nodata & N (188) & \nodata \\
{\bf V598 Pup} & 3.46 (14) & 2454257.79 (14) & 9$\pm1$ (190) & \nodata & 0.16 (190) & 1.4e+21 & \nodata & 2.95$\pm0.8$ (190) & \nodata & \nodata \\
{\bf V679 Car} & 7.55 (191) & 2454797.77 (191) & 20: & \nodata & \nodata & 1.3e+22  & \nodata & \nodata & \nodata & \nodata \\
{\bf V723 Cas} & 7.1 (192) & 2450069.0 (192) & 263 (2) & 600 (193) & 0.5 (194) & 2.35e+21  & 0.69 (195) & 3.86$\pm0.23$ (196) & N & \nodata \\
V838 Her & 5 (197) & 2448340.5 (197) & 2 (198) & \nodata & 0.5$\pm0.1$ (198) & 2.6e+21  & 0.2975 (199) & 3$\pm1$ (198) & Y (200) & \nodata \\
{\bf XMMSL1 J06} & 12 (201) & 2453643.5 (202) & 8$\pm2$ (202) & \nodata & 0.15 (203) & 8.7e+20 & \nodata & 50 & \nodata & \nodata \\
\enddata
\end{deluxetable*}
\end{longrotatetable}

A handy "cheat sheat" that provides the necessary LaTeX to produce 17 
different types of tables is available at \url{http://journals.aas.org/authors/aastex/aasguide.html#table_cheat_sheet}.

%% The reference list follows the main body and any appendices.
%% Use LaTeX's thebibliography environment to mark up your reference list.
%% Note \begin{thebibliography} is followed by an empty set of
%% curly braces.  If you forget this, LaTeX will generate the error
%% "Perhaps a missing \item?".
%%
%% thebibliography produces citations in the text using \bibitem-\cite
%% cross-referencing. Each reference is preceded by a
%% \bibitem command that defines in curly braces the KEY that corresponds
%% to the KEY in the \cite commands (see the first section above).
%% Make sure that you provide a unique KEY for every \bibitem or else the
%% paper will not LaTeX. The square brackets should contain
%% the citation text that LaTeX will insert in
%% place of the \cite commands.

%% We have used macros to produce journal name abbreviations.
%% \aastex provides a number of these for the more frequently-cited journals.
%% See the Author Guide for a list of them.

%% Note that the style of the \bibitem labels (in []) is slightly
%% different from previous examples.  The natbib system solves a host
%% of citation expression problems, but it is necessary to clearly
%% delimit the year from the author name used in the citation.
%% See the natbib documentation for more details and options.

\begin{thebibliography}{}

\bibitem[Astropy Collaboration et al.(2013)]{2013A&A...558A..33A} Astropy Collaboration, Robitaille, T.~P., Tollerud, E.~J., et al.\ 2013, \aap, 558, A33 
\bibitem[Bertin \& Arnouts(1996)]{1996A&AS..117..393B} Bertin, E., \& Arnouts, S.\ 1996, \aaps, 117, 393 
\bibitem[Corrales(2015)]{2015ApJ...805...23C} Corrales, L.\ 2015, \apj, 805, 23
\bibitem[Ferland et al.(2013)]{2013RMxAA..49..137F} Ferland, G.~J., Porter, R.~L., van Hoof, P.~A.~M., et al.\ 2013, \rmxaa, 49, 137
\bibitem[Hanisch \& Biemesderfer(1989)]{1989BAAS...21..780H} Hanisch, R.~J., \& Biemesderfer, C.~D.\ 1989, \baas, 21, 780 
\bibitem[Lamport(1994)]{lamport94} Lamport, L. 1994, LaTeX: A Document Preparation System, 2nd Edition (Boston, Addison-Wesley Professional)
\bibitem[Schwarz et al.(2011)]{2011ApJS..197...31S} Schwarz, G.~J., Ness, J.-U., Osborne, J.~P., et al.\ 2011, \apjs, 197, 31  
\bibitem[Vogt et al.(2014)]{2014ApJ...793..127V} Vogt, F.~P.~A., Dopita, M.~A., Kewley, L.~J., et al.\ 2014, \apj, 793, 127

\end{thebibliography}

%% This command is needed to show the entire author+affilation list when
%% the collaboration and author truncation commands are used.  It has to
%% go at the end of the manuscript.
%\allauthors

%% Include this line if you are using the \added, \replaced, \deleted
%% commands to see a summary list of all changes at the end of the article.
%\listofchanges

\end{document}

% End of file `sample62.tex'.

### 2020-01-21 paper/paper.tex
draftish of paper

@@
 \newcommand{\cf}{2pcf\xspace} %2pF? 2PCF? %TODO: fix spacing after
+% to capitalize or not to capitalize? leaning no
 \newcommand{\Est}{The Continuous-Function Estimator\xspace}
@@
 % Context
-The two-point correlation function (\cf) is the most important statistic in structure formation.
-Current estimators of the \cf, which is used to measure the clustering of density field tracers, have significant limitations.
-The standard Landy-Szalay (\LS) estimator evaluates the \cf in bins of radial separation between objects, and the choice of bins introduces a trade-off between bias and variance.
-%is radial the right word?
-does that imply line-of-sight?
-% Aims
-We present a new, generalized estimator, \est, for the \cf that obviates binning in separation or any other property.
-% Results
-\Est replaces the binned pair counts of \LS with a continuous-function representation in the form of a linear superposition of basis functions.
-These functions can take advantage of the known form of the \cf, and can depend on other properties of the pairs in addition to their separation.
-Inspired by linear least-squares fitting, \est marks the numerator terms of \LS with vectors of these functions, and the denominator with a tensor of their outer products.
-The estimator outputs the best-fit linear combination of basis functions to describe the \cf, in the same limit in which \LS is justifiably optimal for binned applications. 
-% Results
-We demonstrate that it can estimate the clustering of artificial data in representations that provide more accuracy with fewer basis functions than \LS, thus reducing requirements on mocks for covariance estimation.
-We apply \est to the Luminous Red Galaxy sample of the Sloan Digital Sky Survey (SDSS), and show that it can extract a compact set of Fisher-optimal statistics for measuring the location of the baryon acoustic feature.
-We discuss other applications and limitations of \est for present and future studies of large-scale structure, including determining the dependence of clustering on galaxy properties and potentially unifying real-space and Fourier-space approaches by using a Fourier basis representation.
+The two-point correlation function (\cf) is the most important statistic in structure formation, used to measure the clustering of density field tracers (e.g. galaxies).
+Current estimators of the 2pcf, including the standard Landy-Szalay (\LS) estimator, have significant limitations.
+%Aims
+In this work we address the issue that the \LS estimator evaluates the \cf in bins of separation between objects, which results in a loss of information and a trade-off between bias and variance.
+\KSF{there are other aims but they are hard to describe, as they are not ``issues'' with the LS estimator but other capabilities of ours - how to phrase?}
+%Methods
+We present a new estimator for the \cf, \emph{\est}, which generalizes \LS to a continuous representation and obviates binning in separation or any other property. 
+Our estimator replaces the binned pair counts of \LS with a linear superposition of any set of basis functions, and outputs the best-fit linear combination of basis functions to describe the \cf. 
+It is closely related to the information-theory optimal estimator used in linear least-squares fitting.
+The choice of basis can take into account the expected form of the \cf, as well as its dependence on other properties beyond separation.
+%Results
+We show that \est can estimate the clustering of artificial data in representations that provide more accuracy with fewer basis functions than \LS.
+Using a spline basis representation, we show that \est achieves lower bias and lower variance than \LS. 
+We also demonstrate how the estimator can be used to directly estimate the Baryon Acoustic Oscillation scale.
+Critically, these representations reduce the number of mock catalogs required for covariance estimation, a limiting factor in \cf measurements.
+%As a demonstration, we apply the estimator to the Luminous Red Galaxy sample of the Sloan Digital Sky Survey (SDSS). 
+%We show that the continuous-function estimator, with basis functions chosen to be derivatives with respect to the cosmological parameters, provides a direct measurement of cosmological parameters from galaxy observations.
+%Similarly, we use the estimator to directly measure the location of the baryon acoustic feature in the SDSS data.
+We discuss other applications and limitations of \est for present and future studies of large-scale structure, including determining the dependence of clustering on galaxy properties and potentially unifying real-space and Fourier-space approaches to clustering measurements.
+
 \end{abstract}
@@
 %some papers have equations (even this eqn) in intro - but should bump it to motivation and just explain?
+Estimators of the \cf have been studied extensively (\citealt{PeeblesHauser1974}; \citealt{DavisPeebles1983}; \citealt{Hamilton1993}).
 The current standard estimator was proposed by \cite{LandySzalay1993}, hereafter \LS.
@@
 
-These estimators present improvements to \LS, but they all still require binning in separation.
+These estimators present improvements to \LS, but they are still limited to estimates in separation bins.
 Some require additional computational costs or layers of complexity, so the standard formulation of \LS continues to be the default estimator used in most analyses.
@@
 This estimator removes the need for binning and allows for the \cf to be represented by fewer basis functions, requiring fewer mock catalogs to compute the covariance matrix.
-It is particularly well-suited to the analysis of LSS features such as the BAO peak; we find that we can more accurately locate the peak with fewer components.
+It is particularly well-suited to the analysis of LSS features such as the BAO peak; we find that we can accurately locate the peak with fewer components.
 
@@
 This paper is organized as follows. 
-In Section \ref{sec:motiv}, we motivate and derive our estimator. 
-In Section \ref{sec:est}, we prove its correctness and describe our implementation, and demonstrate its application on a simulated data set. 
-Our results on the SDSS LRG sample are shown in Section \ref{sec:app}. 
-We discuss the implications and other possible applications in Section \ref{sec:discuss}.
+In $\S$\ref{sec:motiv}, we motivate our estimator and explain its formulation.
+%In Section \ref{sec:est}, we prove its correctness and describe our implementation, and demonstrate its application on a simulated data set. 
+%Our results on the SDSS LRG sample are shown in Section \ref{sec:app}. 
+We demonstrate its application on a simulated data set, including a toy BAO analysis, in $\S$\ref{sec:app}.
+We discuss the implications and other possible applications in $\S$\ref{sec:discuss}.
 
-\section{Motivation} \label{sec:motiv}
+\section{Motivation and Formulation} 
+\label{sec:motiv}
 
-The pair counts for \LS and related estimators are defined explicitly as
+\subsection{Standard Clustering Estimation}
+
+We assume we have a data catalog with $N_D$ objects and a random catalog with $N_R$ objects.
+The pair counts for \LS and related estimators are then defined explicitly as
 %TODO: fix spacing on sides of equal sign
 \begin{eqnarray}\displaystyle
-DD_k &\equiv& \sum_{n n'} i(g_k < |x_n - x_{n'}| < h_k) \\
-DR_k &\equiv& \sum_{n m} i(g_k < |x_n - x_m| < h_k) \\
-RR_k &\equiv& \sum_{m m'} i(g_k < |x_m - x_{m'}| < h_k),
+DD_k &\equiv& \frac{2}{N_D(N_D-1)} \sum_{n n'} i(g_k < |x_n - x_{n'}| < h_k) \\
+DR_k &\equiv& \frac{1}{N_D N_R} \sum_{n m} i(g_k < |x_n - x_m| < h_k) \\
+RR_k &\equiv& \frac{2}{N_R(N_R-1)} \sum_{m m'} i(g_k < |x_m - x_{m'}| < h_k),
 \end{eqnarray}
-where $DD_k$ is the count of data--data tracer pairs in bin $k$ (which has bin edges $g_k$ and $h_k$), $i()$ is an indicator function, $x$ is the tracer position, the $n$ and $n'$ indices index data positions, the $m$ and $m'$ indices index random catalog positions, $DR_k$ is the count of data--random pairs, and $RR_k$ is the count of random--random pairs.
+where $DD_k$ is the count of data--data tracer pairs in bin $k$ (which has bin edges $g_k$ and $h_k$), $i()$ is an indicator function \KSF{define?}, $x$ is the tracer position, the $n$ and $n'$ indices index data positions, the $m$ and $m'$ indices index random catalog positions, $DR_k$ is the count of data--random pairs, and $RR_k$ is the count of random--random pairs.
+\KSF{Do i need to note that Corrfunc (and some papers) actually kills the -1 in the prefactor? though this might be related to how Corrfunc does pair counting}
 
-The Landy-Szalay estimator is justified by\ldots
+The \LS estimator has been analyzed extensively; it is shown to have the lowest bias and variance compared to other pair-count-based estimators (e.g. \citealt{Kerscher2000}).
+In the limit of unclustered data, for a data volume much larger than the scales of interest, and an infinitely large random catalog, the \LS estimator is known to be optimal: it is unbiased and has minimum variance.
+In practice the latter two limits are sufficiently satisfied, but the data we are interested in is clustered.
+\LS does show a bias on very large scales ($>$150 \hmpc), but the bias is significantly smaller than that of most other estimators (\citealt{Kerscher1999}, \citealt{VargasMagana2013}).
+\LS is also less sensitive to the number of random points than other estimators \citep{Kerscher2000}.
+\cite{VargasMagana2013} show that for clustered data, the \LS estimator has lower variance than other estimators, but does not reach the Poisson noise limit.
+
+\KSF{Should i explain / show some of the other estimators?}
+
+\KSF{Do i need to explain real space vs redshift space vs projected correlation function?}
+
+\subsection{Least Squares Fitting}
 
@@
 % TODO: fix this capitalization issue
-\section{\Est} 
+%\section{\Est} 
+%\label{sec:est}
+
+\subsection{\est}
 \label{sec:est}
@@
 \begin{eqnarray}\displaystyle
-DD &\equiv& \sum_{n n'} f(T_n, T_{n'}) \\
-DR &\equiv& \sum_{n m} f(T_n, T_{m}) \\
-RR &\equiv& \sum_{m m'} f(T_m, T_{m'}) \\
-QQ &\equiv& \sum_{m m'} f(T_m, T_{m'}) \cdot f\T(T_m, T_{m'}),
+DD &\equiv& \frac{2}{N_D(N_D-1)} \sum_{n n'} f(T_n, T_{n'}) \\
+DR &\equiv& \frac{1}{N_D N_R)} \sum_{n m} f(T_n, T_{m}) \\
+RR &\equiv& \frac{2}{N_R(N_R-1)} \sum_{m m'} f(T_m, T_{m'}) \\
+QQ &\equiv& \frac{2}{N_R(N_R-1)} \sum_{m m'} f(T_m, T_{m'}) \cdot f\T(T_m, T_{m'}),
 \end{eqnarray}
@@
 In the case where the basis functions only depend on the pair separation, we have $f = f(|x_n - x_{n'}|)$ and $\xi = \xi(r)$, where $r$ is the separation at which to evaluate the correlation function.
-Specifically, we can choose to define a set of basis functions $f$ as,
+Specifically, we can choose to define a set of basis functions $f^T$ as,
 \begin{equation}
-f_k(|x_n - x_n'|) =  i(g_k < |x_n - x_{n'}| < h_k)
+f^T_k(|x_n - x_n'|) =  i(g_k < |x_n - x_{n'}| < h_k)
 \end{equation}
@@
 In this case the $DD$, $DR$ and $RR$ vectors become binned pair counts and the $QQ$ matrix becomes diagonal, with diagonal elements equal to the $RR$ vector elements. %TODO CONTINUE HERE
-Then for this function $f$, which is the definition of a top-hat (rectangular) function, the estimator is equivalent to that of Landy-Szalay.
+Then for this set of functions $f^T$, which is the definition of a top-hat (rectangular) function, the estimator is equivalent to that of Landy-Szalay.
 
-We prove that the estimator is invariant under affine transformations.
+\KSF{Worth giving analytic formulation of RR term? Or at least mentioning that there is one and we have derived it?}
 
-We show that in the limit of infinitesimal bins\ldots
+\KSF{Discuss the limit of infinitesimal bins? Do we know what this is?}
 
-These show that the estimator is correct.
+We implement this estimator based on the correlation function package \texttt{Corrfunc} by \cite{Sinha2019}.
+The code is open-source and available at \texttt{github.com/kstoreyf/Corrfunc}.
+\KSF{Should give any detail about implementation?}
 
-We implement this estimator based on the correlation function package \texttt{corrfunc} by \citep{citeXXX} \ldots
-The code is open-source and available at XXX.
+\KSF{where to put this link? In a footnote? And is it ok to keep it as a fork?}
 
-\subsection{Demonstration using Spline Basis Functions}
-\label{sec:spline}
+\section{Applications}
+\label{sec:app}
+
+\KSF{Does this count as Applications if its on fake data? Better name?}
 
@@
 
+\subsection{Demonstration using Spline Basis Functions}
+\label{sec:spline}
+
 \label{fig:quadratic}
-\begin{figure}[h]
+\begin{figure}[ht]
 \centering
@@
 Measurement of the BAO scale provides a good use case for our estimator.
+The BAO feature is a peak in clustering on large scales, $\sim$150 Mpc, making it less sensitive to small-scale astrophysical effects.
+It is one of the best tools for constraining cosmological models, in particular the distance-redshift relation (\citealt{Kazin2010}; \citealt{Anderson2012}; \citealt{Anderson2014}; \citealt{Alam2016}).
 
+\KSF{Is there meaning to ``redshift space separation'' when I'm workin in a simulation box? Should I just say separation?}
+
+\KSF{Should this be done on the periodic box? On a box assuming non-periodicity so we need the RR term? On a sky projection of a box, but without a window function?}
+
+\KSF{what else do I need to cite for BAO?}
+
+We base our BAO estimation on the method of the BOSS DR10 and 11 analysis \citep{Anderson2014}.
+We measure the spherically averaged redshift space correlation function, $\xi(s)$, where $s$ is the redshift-space separation between pairs.
 The BAO peak position is a function of multiple distance scales.
@@
 
-For BAO measurements, we measure the spherically averaged redshift space correlation function, $\xi(s)$.
-This multipoles of the correlation function are given by
-\begin{equation}
-\xi_l(s) = \frac{1}{2} \int^{1}_{-1} \xi_l(s, \mu) L_l(\mu) d\mu
-\end{equation}
-where $L_l$ is the Legendre multipole of order $l$,  $s$ is the redshift-space separation between pairs, and $\mu = cos(\theta)$, $\theta$ is the angle between the separation vector $s$ and the line-of-sight direction (i.e.
+% Using spherically averaged so don't need this!!
+% The multipoles of the correlation function are given by
+% \begin{equation}
+% \xi_l(s) = \frac{1}{2} \int^{1}_{-1} \xi_l(s, \mu) L_l(\mu) d\mu
+% \end{equation}
+% where $L_l$ is the Legendre multipole of order $l$,  $s$ is the redshift-space separation between pairs, and $\mu = cos(\theta)$, $\theta$ is the angle between the separation vector $s$ and the line-of-sight direction (i.e.
 $\mu=1$ indicates a galaxy pair along the line of sight, $\mu$=0 indicates a pair with only transverse separation).
-The two-point correlation function that we use is the $l=0$ multiple; from now on we will omit the subscript, $\xi(s) = \xi_0(s)$.
 
@@
 \label{fig:bao_bases}
-\begin{figure}[h]
+\begin{figure}[ht]
 \centering
@@
 \label{fig:bao}
-\begin{figure}[h]
+\begin{figure}[th]
 \centering
@@
 We note that our estimator doesn't rely on \LS; it could be implemented for various linear combination of pair counts, given certain properties. %aka the tensor term
+[Mention other estimators]
 
-We discuss the relation of our estimator to the optimized estimator of \cite{VargasMagana2013}, which also uses a linear combination approach.
+This includes the optimized estimator of \cite{VargasMagana2013}, which also uses a linear combination approach. 
+[Describe how]
+
+\KSF{Mention other work with Barnett about LS estimator?}
 
@@
 
-Our performance is by definition the same (marginally more) compared to traditional estimators due to the limiting factor of pair-finding.
+The computational scaling is by definition the same compared to traditional estimators due to the limiting factor of pair-finding.
+\Est has the additional need for evaluating the function $f$ for each pair of galaxies.
+For simple basis functions like splines, this will only marginally decrease performance.
+For more complicated functions such as evaluating a cosmological model, \Est may incur extra computational expense.
+Basis functions can also be input on a grid and then interpolated; the performance is then the same for all functions, but the interpolation for each function for each pair does somewhat decrease the performance.
 
-\subsection{Computing the Covariance Matrix}
+\KSF{Is there a need to say much more?
+If not, maybe doesn't deserve its own section - could this go in the implementation section?
+Or maybe tacked onto another small section if we have one that makes sense}
 
-The \est results in \cf estimates that are just as accurate with fewer components.
-This is critical when computing the covariance matrix.
-The covariance matrix is used to estimate the error on a measurement, by evaluating the \cf on a large number of mock catalogs and computing the covariance between the bins.
-In our replacement of bins with continuous functions, the covariance matrix will be the covariance between these functions.
-The unbiased estimator for the sample covariance matrix is
+\subsection{Effect on Covariance Matrix Estimation}
+
+We have shown that \est results in \cf estimates that are just as accurate with fewer components.
+This is critical when estimating the covariance matrix, which is necessary for parameter inference.
+The covariance matrix is difficult to compute theoretically;instead, it is usually estimated by evaluating the \cf on a large number of mock catalogs and computing the covariance between the bins (e.g.
+\citealt{Reid2010}; \citealt{Anderson2014}).
+The unbiased estimator for the sample covariance matrix is (e.g. \citealt{Anderson2003})
 \begin{equation}
@@
 The variance in the elements of this estimator then have a dependence on $N_{mocks}$ and $N_{bins}$.
-This propagates to the derived cosmological parameters, resulting in an overestimation of the error bars.
+This propagates to the derived cosmological parameters, resulting in an overestimation of the error bars (\citealt{Hartlap2007}; \citealt{Dodelson2013} \citealt{Percival2014}; \citealt{TaylorJoachimi2014}).
 Assuming that $N_{mocks} >> N_{bins}$ (and both much larger than the number of parameters to be estimated), and that the measurements are Gaussian distribued, the error bars are inflated by a factor of $(1 + N_{bins}/N_{mocks})$ (i.e., the true constraints are tighter than the derived ones).
-This factor is critical at the precision of cosmological parameter estimation.
+This factor becomes critical at the precision of cosmological parameter estimation \citep{Percival2014}.
 
 Typically, this is dealt with by generating a very large number of mocks.
-For the Baryon Oscillation Spectroscopic Survey (BOSS, \citealt{Eisenstein2011}), $\sim$600 mocks were needed and the analysis used 41 bins.
+For the Baryon Oscillation Spectroscopic Survey (BOSS, \citealt{Dawson2013}), $\sim$600 mocks were needed and the analysis used 41 bins \citep{Sanchez2012}.
 Future surveys will have more costly requirements on mock catalogs, with larger simulations necessary to cover the larger survey volumes.
@@
 An alternative to increasing $N_{mocks}$ is decreasing $N_{bins}$ to achieve the same error on precision.
-In the standard method, this is shown to \emph{increase} the statistical error, albeit only slightly \cite{Percival2013}.
+In the standard method, this is shown to \emph{increase} the statistical error, albeit only slightly \cite{Percival2014}.
 A substantial increase in bin width would prevent capturing information in finer clustering features; even the relatively broad BAO peak requires a bin size on the order of its width of $\sim$10\hmpc.
@@
 With our estimator, we have shown that we can reduce the variance by using fewer components, without sacrificing accuracy.
-This means that we can safely reduce $N_{bins}$.
+This means that we can safely reduce $N_{bins}$, or in our replacement of bins with continuous functions, the number of basis functions $K$.
+The covariance matrix will be the covariance between these basis functions. \KSF{worth noting that the structure of this covmat will be significantly different, esp if non-orthogonal?}
 To then achieve the same precision on the error on the cosmological parameters, a lower value of $N_{mocks}$ becomes possible.
@@
 
+\KSF{I think the result of discussions was that there wasn't a good way of showing this without propagating all the way to cosmological parameters.
+Would love a figure showing lower covariance errors but not sure how without full propagation}
+
 \subsection{Further Applications}
 
-In the future, we could simultaneously estimate the correlation function and the power spectrum.
+We could also investigate: gradients, anomalies, growth of structure, directly estimating the cosmological parameters, calibrating for systematics, clustering dependence of galaxy properties.
+[Explain (some of) these in more detail]
 
-We could also investigate: gradients, anomalies, growth of structure, directly estimating the cosmological parameters, calibrating for systematics, clustering dependence galaxy properties
+We also may be able to use our estimator to simultaneously estimate the correlation function and the power spectrum.
+[explain how, implications]
+
+\subsection{Summary}
+
+\KSF{Do we want a summary? Think it might be useful for this particular paper, but def repetitive}
 
@@
 %\bibliographystyle{apj} 
-\bibliography{paper}
+%\bibliography{paper}
+% To copy from mendeley locally: in paper dir, [cp ~/code/bibtex/LSS-est_paper.bib .]
+\bibliography{LSS-est_paper}
 

### 2020-02-07 paper/paper.tex
paper updates - solid draft!

@@
 % comments
-\newcommand{\KSF}[1]{\textcolor{teal}{KSF says: #1}}
+%\newcommand{\KSF}[1]{\textcolor{teal}{KSF says: #1}}
+\newcommand{\KSF}[1]{\textcolor{teal}{}}
 
@@
 In practice the latter two limits are sufficiently satisfied, but the data we are interested in is clustered.
-\LS does show a bias on very large scales ($>$150 \hmpc), but the bias is significantly smaller than that of most other estimators (\citealt{Kerscher1999}, \citealt{VargasMagana2013}).
+\LS does show a bias on very large scales ($>$130 \hmpc), but the bias is significantly smaller than that of most other estimators (\citealt{Kerscher1999}, \citealt{VargasMagana2013}).
 \LS is also less sensitive to the number of random points than other estimators \citep{Kerscher2000}.
@@
 
-\KSF{Should i explain / show some of the other estimators?}
+\KSF{Figure demonstrating trade-off between bias and variance? or too simple}
+
+\KSF{Should i explain / show some of the other estimators? --> now i do in the beyond LS section}
 
@@
 
-\subsection{\est}
+\subsection{\Est}
 \label{sec:est}
@@
 DD &\equiv& \frac{2}{N_D(N_D-1)} \sum_{n n'} f(T_n, T_{n'}) \\
-DR &\equiv& \frac{1}{N_D N_R)} \sum_{n m} f(T_n, T_{m}) \\
+DR &\equiv& \frac{1}{N_D N_R} \sum_{n m} f(T_n, T_{m}) \\
 RR &\equiv& \frac{2}{N_R(N_R-1)} \sum_{m m'} f(T_m, T_{m'}) \\
@@
 % here is where the T notation is weird - when you look at evaluation
-\xi(T_l, T_{l'}) &\equiv& a\T \cdot f(T_l, T_{l'}) \\
-a &\equiv& QQ\inv \cdot (DD + 2DR - RR) 
+a &\equiv& QQ\inv \cdot (DD + 2DR - RR) \\
+\xi(T_l, T_{l'}) &\equiv& a\T \cdot f(T_l, T_{l'})
 \end{eqnarray}
@@
 
-We demonstrate the application of the \est on a simulated Gaussian random field.
-We generate lognormal mock catalogs using the \texttt{nbodykit} package.
+We demonstrate the application of \est on artifical data.
+We generate lognormal mock catalogs \citep{ColesJones1991} using the \texttt{nbodykit} package \citep{Hand2018}.
 We use an input power spectrum with the Planck cosmology \KSF{CITE}.
@@
 Our test catalogs have size (750 \hmpc)$^3$ and a galaxy number density of $3 \times 10^{-4}$, to match that of the Brightest Cluster Galaxies \KSF{CITE}.
-We compute 100 realizations of this box. 
+We construct 100 realizations of this box. 
 We also generate a uniformly sampled random catalog with 10 times the number of galaxies as in the data catalogs.
@@
 
-\label{fig:quadratic}
+\label{fig:spline}
 \begin{figure}[ht]
 \centering
-    \includegraphics[width=0.8\textwidth]{tophat_quadratic}
-    \caption{A comparison between the standard tophat estimator (blue solid) and the \est with a quadratic spline basis function (red dotted).
+    \includegraphics[width=0.8\textwidth]{tophat_cubic}
+    \caption{A comparison between \est with a cubic spline basis function (red dotted) and the standard tophat estimator (blue solid).
 The shaded region is the 1$\sigma$ variation in the 100 mock catalogs.
-The quadratic estimator has lower bias and lower variance with fewer bins.}
+The cubic spline estimator has lower bias and lower variance with fewer bins.
+\KSF{move rmse from figure into text only?
+but then not necessarily clear its less biased} \KSF{does tophat need to be a different linestyle?}}
 \end{figure}
@@
 
-We compare the standard estimator, reformulated as continuous functions using a tophat basis, to the \est with a quadratic spline basis.
-We use second-order B-splines, which constitute the set of basis functions for a quadratic spline \KSF{CITE}.
+A natural extension of tophat basis functions is the B-spline.
+B-splines of order $n$ are piecewise polynomials of order $n-1$; they constitute the basis functions for spline interpolation \KSF{CITE}.
+They have the nice property that the functions and their derivatives can be continuous, depending on the order.
+Further, B-splines are well-localized, which provides a more direct comparison to the typical tophat basis (which is entirely localized).
+For this demonstration we use fourth-order B-splines, which constitute the set of basis functions for a cubic spline, as they are the lowest-order spline to have a continuous first derivative.
+We compare this standard estimator, reformulated as continuous functions using a tophat basis.
 For the tophat basis we use 44 basis functions in the range $40 < r < 150$ \hmpc, each with a width of 2.5 \hmpc. 
-For the quadratic spline basis, we use the same $r$ range, but with only 11 basis functions, and knots chosen on a grid of 10 \hmpc.
+For the cubic spline basis, we use the same $r$ range, but with only 11 basis functions, and knots chosen on a grid of 10 \hmpc.
 \KSF{need to specify more about knots? is this accurate (not quite, the knots repeat values at the edge i think...) and should i mention control points?}
 
-The results are shown in Figure~\ref{fig:quadratic}, compared to the true input \cf.
-The quadratic basis results in a \cf estimate that has a root mean square error (RMSE) with respect to the truth of $4.39 \times 10^{-4}$, compared to $5.43 \times 10^{-4}$ for the tophat basis.
+The results are shown in Figure~\ref{fig:spline}, compared to the true input \cf.
+The spline basis results in a \cf estimate that has a root mean square error (RMSE) with respect to the truth of $4.38 \times 10^{-4}$, compared to $5.31 \times 10^{-4}$ for the tophat basis.
 This holds true when we average over the bin \KSF{need to do weighted average based on which r-values contribute? supposed to hear from tinker}, as in standard practice, and then compute the RMSE.
-The quadratic basis also results in lower variance across the 100 mocks, compared to the tophat basis.
-Thus, the \est allows for a more accurate and precise estimate of the \cf with fewer components.
+The spline basis also results in lower variance across the 100 mocks, compared to the tophat basis.
+Thus, \est allows for a more accurate and precise estimate of the \cf with fewer components.
+
+We note that this is fundamentally different than a kernel-based estimator.
+In a kernel formulation, each data point is smoothed by a particular kernel, smoothing out features in the resulting function.
+With our estimator, the basis functions are fixed and the data is projected onto them.
+This preserves the information in the data to the degree given by the chosen set of basis functions, which can in fact enhance features rather than smooth them.
+\KSF{please add/correct here}
 
@@
 The BAO feature is a peak in clustering on large scales, $\sim$150 Mpc, making it less sensitive to small-scale astrophysical effects.
-It is one of the best tools for constraining cosmological models, in particular the distance-redshift relation (\citealt{Kazin2010}; \citealt{Anderson2012}; \citealt{Anderson2014}; \citealt{Alam2016}).
+It is one of the best tools for constraining cosmological models, in particular the distance-redshift relation \citep{Kazin2010, Anderson2012, Anderson2014, Alam2016}.
 
@@
 where $D_A$ is the angular diameter distance, $H$ is the Hubble constant, $r_s$ is the sound horizon scale at the drag epoch, and the superscript ``$mod$'' denotes the value for the fiducial model.
+\KSF{mod or fid?
+fid just was v close to fit}
 If we find $\alpha>1$, the peak is shifted to smaller scales than its position in the model, and if $\alpha<1$, the peak is shifted to larger scales.
@@
 The resulting value for $\alpha$ is used to derive the actual values of the distance scales of interest.
+Typically, density-field reconstruction is performed before applying the estimator to correct for nonlinear growth around the BAO scale \citep{Eisenstein2007}; for our toy example, we will omit this step.
 
@@
 It converges to the same, but wrong, value of alpha, and diff ones for diff realizations.
-Need to average over many?
-(Still maybe wrong due to some persistant weirdness as mentioned above!
-Maybe need to check how I'm doing the lognormals)}
+Need to average over many?}
+
+\KSF{compare to tophat fit with standard BAO fitting function??}
 
@@
 The $a_1$, $a_2$, and $a_3$ terms are nuisance parameters to fit the broadband shape.
-The \KSF{Make colorblind friendly!
-Do i also need to use different linestyles?
-Or maybe weights?}}
+\KSF{Make colorblind friendly!
+Do i also need to use different linestyles?}}
 \end{figure}
@@
 We then take that value as our $\alpha_{mod}$, and re-run the estimator.
-We stop when the percent change between $\alpha_{mod}$ and the recovered $\alpha$ is less than 0.01\%; it typically takes only $\sim$5 iterations to converge.
+We stop when the percent change between $\alpha_{mod}$ and the recovered $\alpha$ is less than 0.01\%; it typically takes only $\sim$5 iterations to converge. \KSF{update this with true values}
 The resulting estimate for the correlation function is shown in Figure~\ref{fig:bao}, in comparison to the traditional tophat estimator.
@@
 
+These basis functions are significantly different than the tophat or B-spline bases previously explored as they are not localized.
+This means that data at all scales could contribute to all basis functions.
+It is then critical to ensure that the final parameter estimate does not reply on the range of scales chosen.
+We have confirmed that in this application, the result is robust to the chosen range as long as the scales cover the range $40<r<200$ \hmpc, the typical range used in BAO analyses.
+\KSF{Actually have to check this and update numbers!}
+
 % \section{Application to survey data} \label{sec:app}
@@
 
-We estimated the \cf with fewer components to achieve the same level of precision.
-
 \subsection{Beyond the Landy-Szalay Estimator}
 
-We note that our estimator doesn't rely on \LS; it could be implemented for various linear combination of pair counts, given certain properties. %aka the tensor term
-[Mention other estimators]
-
-This includes the optimized estimator of \cite{VargasMagana2013}, which also uses a linear combination approach. 
-[Describe how]
+While we have formulated our estimator as a generalization of \LS, as it is the standard used in \cf analyses and has optimal properties under certain conditions, we can also reformulate it for other estimators.
+The formulation currently requires a normalization term (i.e. denominator) of the $RR$ counts, as we replace this with our $QQ$ term.
+This is the case for the \cite{PeeblesHauser1974} estimator and the \cite{Hewett1982} estimator:
+\begin{eqnarray}
+    \xi_{PH}(r) &=& \frac{DD - RR}{RR} \\
+    \xi_{Hew}(r) &=& \frac{DD - DR}{RR}.
+\end{eqnarray}
+\KSF{dropped $k$ and $(r)$ notation here, do i need?}
+We could also generalize estimators which have a $DR$ term as the denominator, such as the \cite{DavisPeebles1983} estimator,
+\begin{equation}
+    \xi_{DP}(r) = \frac{DD - DR}{DR}
+\end{equation}
+by defining
+\begin{equation}
+    DQ = \frac{2}{N_D N_R} \sum_{n m} f(T_n, T_m) \cdot f\T(T_n, T_m).
+\end{equation}
+This could be extended to almost any linear combination of pair counts.
+The estimator of \cite{VargasMagana2013} selects the optimal combination of pair counts; our estimators could be combined to create an even more generalized estimator.
+\KSF{some of the terms in V-M include DD in the denom - need to mention this?}
 
@@
 For simple basis functions like splines, this will only marginally decrease performance.
-For more complicated functions such as evaluating a cosmological model, \Est may incur extra computational expense.
+For more complicated functions such as evaluating a cosmological model, \est may incur extra computational expense.
 Basis functions can also be input on a grid and then interpolated; the performance is then the same for all functions, but the interpolation for each function for each pair does somewhat decrease the performance.
@@
 
-We could also investigate: gradients, anomalies, growth of structure, directly estimating the cosmological parameters, calibrating for systematics, clustering dependence of galaxy properties.
-[Explain (some of) these in more detail]
+The formulation of \est opens up many possibilities for extracting information from the correlation function.
+The most straightforward applications are standard basis functions or linearizeable astrophysical models, as we have shown here.
+Other applications for the direct estimation of cosmological parameters could include the growth rate of cosmic structure $f$ \citep{Satpathy2016, Reid2018} and primordial non-Gaussianity in the local density field $f^{local}_{NL}$ \citep{Karagiannis2014}.
+\KSF{mention idea of doing full cosmo model analysis by taking derivs wrt cosmological params? cool but less connected to citeable papers perhaps}
 
-We also may be able to use our estimator to simultaneously estimate the correlation function and the power spectrum.
-[explain how, implications]
+We can take our estimator a step further by choosing basis functions that depend not only on the separation between tracer pairs, but also on the properties of the tracers themselves.
+One such application is the redshift dependence of the Alcock-Paczynski effect \cite{AlcockPaczynski1979}, which can be used to constrain the matter density $\Omega_m$ and the dark energy equation of state parameter $w$ \citep{Li2016}.
+The basis functions $f$ in this case would take the form
+\begin{equation}
+    f_k(T_n, T_m) = f_k(r_{nm}, z_n, z_m),
+\end{equation}
+where $z$ is the redshift of tracer $n$ or $m$.
+Another potential use case is the luminosity and color dependence of galaxy clustering, which can be used to probe galaxy formation \citep{Zehavi20}.
+This could be extended to other galaxy properties.
 
-\subsection{Summary}
+The estimator gives us the opportunity to investigate more subtle or exotic signals which are anomalous with respect to our conventional models.
+Anomalies could appear as inhomogeneities or anisotropies in the data.
+For example, cosmological parameters could vary across the sky, which has previously investigated in patches across the Cosmic Microwave Background \citep{MukherjeeWandelt2018}.
+Another possibility is anisotropy in the cosmic acceleration, which could appear in various probes including the BAO peak \citep{Faltenbacher2012} and Type Ia supernovae \citep{Colin2019}.
+With our estimator, we could introduce a dependence on location or direction into our basis functions, and constrain the potential deviation from homogeneity or isotropy.
+While these effects would be highly degenerate with systematics, our estimator combined with robust systematics mitigation allows us to investigate the possibility of new physics.
+
+Finally, our estimator can be directly related to a power spectrum analysis.
+We could use a Fourier basis as our set of continuous functions.
+This would allow us to directly project the data onto Fourier modes.
+This represents a step towards unifying the correlation function and the power spectrum. \KSF{there's more to say here but I'm not sure what}
+
+\section{Summary}
+
+We did cool science!
 

### 2020-02-11 paper/paper.tex
updated bao figure

@@
 % -----
-% - get yourself an ORCID and use it.
-% - Is ``vectorized'' the right word? It's what I say, but now I'm doubting it. We could go with linear-regression. Or affine-invariant. Or unbinned. Or continuous-function?
 
@@
 % -----------
-% - line break at sentence breaks? Or more frequently? So git diff works good?
+% - line break at sentence breaks
 
@@
 % comments
-%\newcommand{\KSF}[1]{\textcolor{teal}{KSF says: #1}}
-\newcommand{\KSF}[1]{\textcolor{teal}{}}
+\newcommand{\KSF}[1]{\textcolor{teal}{KSF says: #1}}
+%\newcommand{\KSF}[1]{\textcolor{teal}{}}
 
@@
 Upcoming galaxy surveys will observe larger volumes with improved measurements, allowing us to test $\Lambda$CDM to even higher precision.
+\KSF{Should I name-drop surveys, and their data volume / sky area?}
 
@@
 These tracers are often to taken to be galaxies, but can also be galaxy clusters, quasars and other sources.
-% should i say that hereafter we will take them to be galaxies?
-The clustering of these objects is measured with two-point statistics, namely the power spectrum $P(k)$ and the two-point correlation function (\cf). %name xi_r?
+\KSF{should i say that hereafter we will take them to be galaxies?}
+The clustering of these objects is measured with two-point statistics, namely the power spectrum $P(k)$ and the two-point correlation function (\cf).
 These characterize the clustering in Fourier space and real space, respectively, with the \cf defined as the Fourier Transform of the power spectrum: 
-%TODO: CHECK, PROBS WRONG
 \begin{equation}
-\xi(r) = \frac{1}{2\pi} \int_0^{\infty} P(k) d^3k
+\xi(r) = \frac{1}{2\pi^2} \int_0^{\infty} P(k) \frac{\mathrm{sin}(kr)}{kr} k^2 dk
 \end{equation}
+where we assume isotropy. \KSF{should i give the original Fourier transform before itegrating over orientation?}
 In principle, the power spectrum and the two-point correlation function contain the same information.
@@
 To account for the survey boundaries as well as corrupted regions due to issues such as bright foreground stars, a set of random points are Poisson-distributed within the acceptable survey window. 
-The pairwise correlations of these points, which by construction are unclustered, are used to normalize out the survey window when estimating the \cf of the clustered data.
+The pairwise correlations of these unclustered points are used to normalize out the survey window when estimating the \cf of the clustered data.
 Typically, this requires random points on the order of 10-100 times the number of data points, making the random correlations the limiting factor in \cf computation.
@@
 The \cf is computed in bins of radial separation, meaning that in practice it measures the volume average of the \cf over the bin. %is this accurate?
-The choice of bins requires a trade-off between bias and variance: fewer bins may bias the result, while more bins increases the variance of measurement.
+This binning introduces inherent limitations.
+First, the choice of bins requires a trade-off between bias and variance: fewer bins may bias the result, while more bins increases the variance of measurement.
 Finite-width bins also result in a loss of information about the property in which one is binning.
-Especially as we strive for extreme precision in large-scale structure analyses, we should be maximizing the information we extract from the data.
-Moreover, we note that standard estimators are fundamentally biased due to this finite binning; we will address this in a future work.
+As we strive for extreme precision in large-scale structure analyses, we should be maximizing the information we extract from the data.
+
+More generally, binning adds arbitrary boundaries between continuous data; results should not depend on bin choice, yet they sometimes do.
+\cite{Lanzuisi2017} noted that the choice of binning axis impacts the detected correlation between the luminosity of active galactic nuclei and their host galaxies; \cite{Grimmett2020} devised a method to investigate this correlation in a continuous manner using a hierarchical Bayesian model, eliminating the need for binning.
+\cite{Bailoni2016} explored the dependence of clustering analyses on the number of redshift bins, finding a non-negligible difference in cosmological parameter uncertainties.
+The implications for BAO analyses were explored by \citep{Percival2014}, who found that there is an optimal bin width given the analysis method.
+This balances the increasing statistical error with small bins and the offset in the derived BAO peak location with large bins; the effects are small but non-negligible.
+%Moreover, we note that standard estimators are fundamentally biased due to this finite binning; we will address this in a future work.
 %TODO: figure out this wording.
@@
 The marked correlation function \citep{WhitePadmanabhan2009} sums weights based on properties of the galaxies one is interested in, such as the local density or galaxy luminosity, and avoids the use of a random catalog.
-The estimators described so far have all taken probabilistic approaches; some have also taken a likelihood approach.
+The estimators described so far have all taken probabilistic approaches; others have taken a likelihood approach.
 \cite{BaxterRozo2013} introduced a maximum likelihood estimator for the \cf, which achieves lower variance compared to the \LS estimator, enabling finer binning and requiring a smaller random catalog for the same precision.
@@
 Some require additional computational costs or layers of complexity, so the standard formulation of \LS continues to be the default estimator used in most analyses.
-%make sure this is true!
-talk to cosmologists who do these analyses
 
@@
 From this connection, we can infer the form of the estimator.
+\KSF{expand this section}
 
 % TODO: fix this capitalization issue
-%\section{\Est} 
-%\label{sec:est}
-
 \subsection{\Est}
@@
 The cubic spline estimator has lower bias and lower variance with fewer bins.
-\KSF{move rmse from figure into text only?
-but then not necessarily clear its less biased} \KSF{does tophat need to be a different linestyle?}}
+\KSF{quote rmse in caption?
+otherwise not clear from figure its less biased.} \KSF{does tophat need to be a different linestyle?}} \KSF{fix to dotted linestyle}
+
 \end{figure}
@@
 \KSF{FINAL PARAMS:  number density 3e-4 (now 1e-4). rmax 200 hMpc (now 150). rcont=1000 (now rcont=300). L=1000 hMpc (now 750 - need more to account for higher rmax)}
-
-\KSF{should RMSE be shown in figure as well as described in text?}
 
@@
 
-\KSF{BUT my estimator is having a hard time recovering the truth anyway!
-Because of variance in the realizations.
-It converges to the same, but wrong, value of alpha, and diff ones for diff realizations.
-Need to average over many?}
-
 \KSF{compare to tophat fit with standard BAO fitting function??}
@@
 We demonstrate this method using the same set of lognormal mocks as in $\S$\ref{sec:spline}.
-We construct a simple test in which our model is the known input model, the Planck parameters, but with some scale dilation offset, here $\alpha_{mod}=1.05$ (while the true value is, of course, $\alpha=1$).
+We construct a simple test in which our model is the known input model, the Planck parameters, but with some scale dilation offset, here $\alpha_{mod}=0.98$ (while the true value is, of course, $\alpha=1$).
 The set of basis functions for our estimator is shown in Figure~\ref{fig:bao_bases}.
@@
 We perform an iterative procedure to estimate $\alpha$.
-After the first estimation with the $\alpha=1.05$, model, we use the value of $C$ to compute the recovered $\alpha$, as described above.
+After the first estimation with the $\alpha=0.98$, model, we use the value of $C$ to compute the recovered $\alpha$, as described above.
 We then take that value as our $\alpha_{mod}$, and re-run the estimator.
 We stop when the percent change between $\alpha_{mod}$ and the recovered $\alpha$ is less than 0.01\%; it typically takes only $\sim$5 iterations to converge. \KSF{update this with true values}
-The resulting estimate for the correlation function is shown in Figure~\ref{fig:bao}, in comparison to the traditional tophat estimator.
+We apply this procedure to each of the 100 mocks; the resulting estimate for the correlation function is shown in Figure~\ref{fig:bao}.
+The mean BAO estimate is shown in orange, and the mean tophat estimate is in blue; the truth is in black.
+Our estimator clearly better represents the shape of the known \cf.
+The mean value of the final recovered scale dilation parameter is $\alpha=0.9995$ \KSF{What do i quote for the error here?}, very close \KSF{quote percent?} to the true value of $1$.
 
@@
     \includegraphics[width=0.8\textwidth]{bao}
-    \caption{Estimation of the correlation function using our estimator with basis functions based on the BAO fitting function (purple dot-dashed).
+    \caption{Estimation of the correlation function using our estimator with basis functions based on the BAO fitting function (orange dot-dashed).
+The line is the mean of the final estimate from the iteration procedure for 100 mocks, and the shaded region is the $1\sigma$ variation.
 We also show the standard Landy-Szalay estimator, displayed as a tophat function (blue), as well as the true input correlation function (black).
-\KSF{This is a placeholder figure!
-Need to resolve issues above.} \KSF{Should this and the basis set figure be a single figure with two panels?}}
+\KSF{This is for the 1e-4 box, need to run bigger box.} \KSF{Should this and the basis set figure be a single figure with two panels?} \KSF{Should i have a figure showing the sum of the basis functions and how that gets you the final cf?}}
 \end{figure}
 
-These basis functions are significantly different than the tophat or B-spline bases previously explored as they are not localized.
+We note that these basis functions are significantly different than the tophat or B-spline bases previously explored, mainly because they are not localized.
 This means that data at all scales could contribute to all basis functions.
-It is then critical to ensure that the final parameter estimate does not reply on the range of scales chosen.
+It is then critical to ensure that the final parameter estimate does not rely on the range of scales chosen.
 We have confirmed that in this application, the result is robust to the chosen range as long as the scales cover the range $40<r<200$ \hmpc, the typical range used in BAO analyses.
@@
 where $z$ is the redshift of tracer $n$ or $m$.
-Another potential use case is the luminosity and color dependence of galaxy clustering, which can be used to probe galaxy formation \citep{Zehavi20}.
+Another potential use case is the luminosity and color dependence of galaxy clustering, which can be used to probe galaxy formation \citep{Zehavi2011}.
 This could be extended to other galaxy properties.

### 2020-05-05 paper/paper.tex
paper tweaks
hogg responses to ksf comments

@@
 \newcommand{\KSF}[1]{\textcolor{teal}{KSF says: #1}}
+\newcommand{\hogg}[1]{\textcolor{teal}{Hogg says: #1}}
 %\newcommand{\KSF}[1]{\textcolor{teal}{}}
@@
 %Aims
-In this work we address the issue that the \LS estimator evaluates the \cf in bins of separation between objects, which results in a loss of information and a trade-off between bias and variance.
-\KSF{there are other aims but they are hard to describe, as they are not ``issues'' with the LS estimator but other capabilities of ours - how to phrase?}
+In this work we address the issue that the \LS estimator evaluates the \cf in bins of separation between objects, which results in a loss of information and a poor trade-off between bias and variance, given the inappropriateness of hard bin edges in most scientific contexts.
+\hogg{The issue is NOT the bias--variance trade-off; that exists everywhere.
+It is that this trade-off is bad when the model is inappropriate.
+It is the inappropriateness that we are addressing.
+I still don't like this wording, but I tried.}
 %Methods
@@
 Upcoming galaxy surveys will observe larger volumes with improved measurements, allowing us to test $\Lambda$CDM to even higher precision.
-\KSF{Should I name-drop surveys, and their data volume / sky area?}
+\KSF{Should I name-drop surveys, and their data volume / sky area?} \hogg{I don't think it is necessary; use your judgement.}
 
@@
 These tracers are often to taken to be galaxies, but can also be galaxy clusters, quasars and other sources.
-\KSF{should i say that hereafter we will take them to be galaxies?}
+\KSF{should i say that hereafter we will take them to be galaxies?} \hogg{Yes.}
 The clustering of these objects is measured with two-point statistics, namely the power spectrum $P(k)$ and the two-point correlation function (\cf).
@@
 \end{equation}
-where we assume isotropy. \KSF{should i give the original Fourier transform before itegrating over orientation?}
+where we assume isotropy. \KSF{should i give the original Fourier transform before itegrating over orientation?} \hogg{Yes, and say that the second follows from the first.}
 In principle, the power spectrum and the two-point correlation function contain the same information.
@@
 \end{eqnarray}
-where $DD_k$ is the count of data--data tracer pairs in bin $k$ (which has bin edges $g_k$ and $h_k$), $i()$ is an indicator function \KSF{define?}, $x$ is the tracer position, the $n$ and $n'$ indices index data positions, the $m$ and $m'$ indices index random catalog positions, $DR_k$ is the count of data--random pairs, and $RR_k$ is the count of random--random pairs.
-\KSF{Do i need to note that Corrfunc (and some papers) actually kills the -1 in the prefactor? though this might be related to how Corrfunc does pair counting}
+where $DD_k$ is the count of data--data tracer pairs in bin $k$ (which has bin edges $g_k$ and $h_k$), $i()$ is an indicator function \KSF{define?} \hogg{yes}, $x$ is the tracer position, the $n$ and $n'$ indices index data positions, the $m$ and $m'$ indices index random catalog positions, $DR_k$ is the count of data--random pairs, and $RR_k$ is the count of random--random pairs.
+\KSF{Do i need to note that Corrfunc (and some papers) actually kills the -1 in the prefactor?
+though this might be related to how Corrfunc does pair counting} \hogg{I have no idea what you are asking here.}
 
@@
 
-\KSF{Figure demonstrating trade-off between bias and variance? or too simple}
+\KSF{Figure demonstrating trade-off between bias and variance? or too simple} \hogg{Only if the point isn't clear without it.}
 
-\KSF{Should i explain / show some of the other estimators? --> now i do in the beyond LS section}
+\KSF{Should i explain / show some of the other estimators?
+--> now i do in the beyond LS section} \hogg{No; our job is not to work through all estimators.
+But we SHOULD note that our mod can be applied to almost any estimator that currently uses bins!}
 
-\KSF{Do i need to explain real space vs redshift space vs projected correlation function?}
+\KSF{Do i need to explain real space vs redshift space vs projected correlation function?} \hogg{No, but always be absolutely unambiguous about what space you are in.
+To the point that you might even want to use different symbols.}
 
@@
 
+% there's a reason im not using k here! because these need to be vectors / tensors to then compute the amplitude
 We generalize the \LS estimator defined above to any set of $K$ basis functions $f$ of the pair, so we now have
@@
 
-\KSF{Worth giving analytic formulation of RR term? Or at least mentioning that there is one and we have derived it?}
+\KSF{Worth giving analytic formulation of RR term? Or at least mentioning that there is one and we have derived it?} \hogg{Huh? We have one?}
 
-\KSF{Discuss the limit of infinitesimal bins? Do we know what this is?}
+\KSF{Discuss the limit of infinitesimal bins? Do we know what this is?} \hogg{Yes, we should discuss this, and show how we are related to that.}
 
@@
 The code is open-source and available at \texttt{github.com/kstoreyf/Corrfunc}.
-\KSF{Should give any detail about implementation?}
+\KSF{Should give any detail about implementation?} \hogg{Yes, we should summarize why this is a good package to build on.}
 
-\KSF{where to put this link? In a footnote? And is it ok to keep it as a fork?}
+\KSF{where to put this link?
+In a footnote?
+And is it ok to keep it as a fork?} \hogg{Put link here and in a note at the end of the acks at the end which says ``all of the code is available open-source at...}
 
@@
 
-\KSF{Does this count as Applications if its on fake data? Better name?}
+\KSF{Does this count as Applications if its on fake data? Better name?} \hogg{I say ``Experiments and results'' these days.}
 
@@
 \KSF{quote rmse in caption?
-otherwise not clear from figure its less biased.} \KSF{does tophat need to be a different linestyle?}} \KSF{fix to dotted linestyle}
+otherwise not clear from figure its less biased.} \hogg{yes} \KSF{does tophat need to be a different linestyle?}} \KSF{fix to dotted linestyle}
 
@@
 
-\KSF{Should i have a figure showing the basis functions?}
+\KSF{Should i have a figure showing the basis functions?} \hogg{yes}
 

### 2020-06-04 paper/paper.tex
red
starting to address hogg edits

@@
 \newcommand{\KSF}[1]{\textcolor{teal}{KSF says: #1}}
-\newcommand{\hogg}[1]{\textcolor{teal}{Hogg says: #1}}
+\newcommand{\hogg}[1]{\textcolor{red}{Hogg says: #1}}
 %\newcommand{\KSF}[1]{\textcolor{teal}{}}
@@
 These characterize the clustering in Fourier space and real space, respectively, with the \cf defined as the Fourier Transform of the power spectrum: 
+\KSF{i'm not sure about my notation here with the vectors and d3k}
 \begin{equation}
-\xi(r) = \frac{1}{2\pi^2} \int_0^{\infty} P(k) \frac{\mathrm{sin}(kr)}{kr} k^2 dk
+\xi(\mathbf{r}) = \frac{1}{(2\pi)^3} \int P(\mathbf{k}) e^{i\mathbf{k} \cdot \mathbf{r}} d^3\mathbf{k}.
 \end{equation}
-where we assume isotropy. \KSF{should i give the original Fourier transform before itegrating over orientation?} \hogg{Yes, and say that the second follows from the first.}
+If we assume isotropy, we find that the spherically averaged correlation function is
+\begin{equation}
+\xi(r) = \frac{1}{2\pi^2} \int_0^{\infty} P(k) \frac{\mathrm{sin}(kr)}{kr} k^2 dk.
+\end{equation}
 In principle, the power spectrum and the two-point correlation function contain the same information.

### 2020-06-04 paper/paper.tex
Added section on analytic computation of RR and QQ

@@
 RR &\equiv& \frac{2}{N_R(N_R-1)} \sum_{m m'} f(T_m, T_{m'}) \\
-QQ &\equiv& \frac{2}{N_R(N_R-1)} \sum_{m m'} f(T_m, T_{m'}) \cdot f\T(T_m, T_{m'}),
+QQ &\equiv& \frac{2}{N_R(N_R-1)} \sum_{m m'} f(T_m, T_{m'}) \cdot f\T(T_m, T_{m'}), \label{eq:qq_proj}
 \end{eqnarray}
@@
 a &\equiv& QQ\inv \cdot (DD + 2DR - RR) \\
-\xi(T_l, T_{l'}) &\equiv& a\T \cdot f(T_l, T_{l'})
+\xi(T_l, T_{l'}) &\equiv& a\T \cdot f(T_l, T_{l'}) \label{eq:xi_proj}
 \end{eqnarray}
@@
 
+\section{Computing RR and QQ Analytically}\label{sec:analytic}
+
+The autocorrelation of the random catalog is meant to approximate the window function. 
+When we have a periodic cube, we can compute this $RR$ term analytically.
+Here we derive this, and then derive the equivalent for our continuous-basis $RR$ and $QQ$ terms.
+
+We consider an annulus around a single galaxy. 
+This annulus has a volume $V_{ann}$.
+Taking the box to have average number density $\bar{n}$, the number of galaxies expected in the annulus is $N_{ann} = A \bar{n}$, and thus this galaxy is part of $N_{ann}$ pairs.   
+We do this for each of the $N_D-1$ other galaxies, and find a total of $(N_D-1) N_{annulus} = \frac{1}{2} (N_D-1) V_{ann} \bar{n}$ pairs, where the factor of $\frac{1}{2}$ accounts for the fact that this double-counts pairs.
+For a cube, $\bar{n} = \frac{N_D}{L^3}$, so we finally count $\frac{1}{2} \frac{N_D}{L^3} (N-1) V_{ann}$ pairs.
+
+For hard-edged radial bins, we can compute $A$ simply as the difference between spherical volumes. 
+We can also represent this as an integral:
+\begin{equation} \label{eq:vol_tophat}
+V_{ann} = \int_{b_1}^{b_2} dV = 4\pi \int_{b_1}^{b_2} r^2 dr
+\end{equation} 
+We can generalize this to any basis function $f(r)$ that is a function of $r$:
+\begin{equation}
+V_{ann} = 4\pi  \int_{b_1}^{b_2} f^2(r) r^2 dr
+\end{equation}
+which we can see reduces to Equation \ref{eq:vol_tophat} when $f(r)$ is the tophat function (returning 1 or 0 depending on whether $r$ is between $b_1$ and $b_2$).
+
+This gives us our full generalized analytic $RR$ term, which has elements
+\begin{equation}
+RR_{i,\text{ana}} = \frac{1}{2} \frac{N_D}{L^3} (N_D-1)  4\pi  \int_{0}^{r_\mathrm{max}} f_i^2(r) r^2 dr
+\end{equation}
+where $i$ is the index of the basis function vector.
+Based on the definition of $QQ$ in Equation \ref{eq:qq_proj} as the outer product of the basis function vector and its transpose, we can see that the elements analytic $QQ$ term are:
+\begin{equation}
+QQ_{ij,\text{ana}} = \frac{1}{2} \frac{N_D}{L^3} (N_D-1)  4\pi  \int_{0}^{r_\mathrm{max}} f_i(r) f_j(r) r^2 dr
+\end{equation}
+This could be further generalized to account for basis functions that take other properties as input.
+
+When considering a periodic box, the naive estimator is no longer biased, so we can also avoid computing the $DR$ term and calculate the amplitudes as 
+\begin{equation}
+a_{\mathrm{ana}} = QQ_\mathrm{ana} \inv \cdot DD.
+\end{equation}
+Looking back, it might have seemed strange that we use $N_D$ in calculating the analytical $RR$ term, but we now see that this normalization prefactor cancels out with that of the $DD$ term.
+Finally, we can compute the correlation function as before in Equation \ref{eq:xi_proj}.
+
 %need style file error
